### 2022-02-07 src/docs_01-07-21-stats-epi-group/01-07-21-stats-epi-group.Rmd
Add the presentation I gave last summer about this work (to-do: use depends for the epil results now that everything is together here)

---
title: Eventually an introduction to the `aghq` R package
subtitle: Statistics and Epidemiology Group Meeting
author: Adam Howes
institute: Imperial College London
date: July 2021
bibliography: citations.bib
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7 
    fig_height: 3
    includes:
        in_header: preamble.tex
---

# Motivation I

* People on ART $A_i$ can be used as supplementary data for small-area estimation of HIV prevalence $\rho_i$
\begin{align*}
A_i &\sim \text{Bin}(N_i, \rho_i \alpha_i), \\
y_i &\sim \text{Bin}(m_i, \rho_i), \\
\text{logit}(\alpha_i) &\sim f(\vartheta_\alpha), \\
\text{logit}(\rho_i) &\sim g(\vartheta_\rho), \quad i = 1, \ldots, n,
\end{align*}
* If $f$ and $g$ are Gaussian then model is almost, but not quite, a latent Gaussian model by the definition of @rue2009approximate
  * This is due to small non-linearities (multiplying two latent Gaussian fields)
  * Each observation depends on more than one element of the latent field

# Motivation II

:::::::::::::: {.columns}

::: {.column width=.65}

* Previous slide is a simplified version of the Naomi evidence synthesis model
* Countries to fit the model using their own data ("in production"?)
  * Can't run long MCMC on the cluster for weeks, as might be the case if this was one paper
* Can't use `R-INLA`, require something more flexible
* Currently using Template Model Builder `TMB` [@kristensen2015tmb]

:::

::: {.column width=.35}

```{r, echo=FALSE, fig.cap="A supermodel", out.width = '65%'}
knitr::include_graphics("naomi_hex.png")
```

:::

::::::::::::::

# Aside: common theme I

* Combining flawed (sparse, aggregated) gold standard (measuring the thing we want) data with other correlated (more available, high resolution) data (measuring not exactly what we want)
* Consistently resulting in models with multiple outcomes (evidence synthesis, multi-output)
* I think a lot of these are going to be not quite LGMs

# Aside: common theme II

* Examples include
  * Naomi model: DHS data is "gold standard", supported by ANC data from pregnant women
  * Sexual risk behaviour model: estimates of FSW population at national level, supported by DHS data
    * The national-level FSW estimates might be more like "bronze standard"
    * DHS approximately asks "have you received money or gifts in exchange for sex in past 12 months"
  * Loa loa prevalence and eyeworm history prevalence model: measuring eyeworm history is a cheap proxy for Loa loa [@amoah2020geostatistical]

# Recap on latent Gaussian models

* Three-stage model
\begin{alignat*}{2}
  &\text{(Observations)}     &   \y &\sim p(\y \, | \, \x), \\
  &\text{(Latent field)}     &   \x &\sim p(\x \, | \, \btheta), \\
  &\text{(Hyperparameters)}  &   \qquad \btheta &\sim p(\btheta),
\end{alignat*}
where $\y = (y_1, \ldots, y_n)$, $\x = (x_1, \ldots, x_n)$, $\btheta = (\theta_1, \ldots, \theta_m)$
* Interested in learning both $(\btheta, \x)$ from data $\y$
* Covers most of the models used in spatiotemporal statistics

# Recap on Integrated Nested Laplace Approximation I

* @rue2009approximate or e.g. @blangiardo2015spatial
* Approximate Bayesian inference for \textcolor{hilit}{latent Gaussian models} (LGMs), which are three-stage models with middle layer
\begin{equation*}
  \text{(Latent field)} \qquad  p(\x \, | \, \btheta) = \mathcal{N}(\x  \, | \, \bm{\mu}(\btheta), \bm{Q}(\btheta)^{-1}).
\end{equation*}
* `R-INLA` implementation takes advantage of sparsity properties of $\bm{Q}(\btheta)$, i.e. if $\x$ is a Gaussian Markov random field (GMRF)

# Recap on Integrated Nested Laplace Approximation II

* Gives approximate \textcolor{hilit}{posterior marginals} $\{\tilde p(x_i \, | \, \y)\}_{i = 1}^n$ and $\{\tilde p(\theta_j \, | \, \y)\}_{j = 1}^m$
* To approximate posterior marginals below requires $\tilde p(\btheta \, | \, \y)$ and $\tilde p(x_i \, | \, \btheta, \y)$
\begin{align}
  p(x_i \, | \, \y) &= \int p(x_i, \btheta \, | \, \y) \text{d} \btheta = \int p(x_i \, | \, \btheta, \y) p(\btheta \, | \, \y) \text{d}\btheta, \quad i = 1, \dots, n, \label{eq:inla1} \\
  p(\theta_j \, | \, \y) &= \int p(\btheta \, | \, \y) \text{d}\btheta_{-j} \quad j = 1, \ldots, m. \label{eq:inla2}
\end{align}

# Recap on Integrated Nested Laplace Approximation III

1) First Laplace approximate hyperparameter posterior
\begin{equation}
\tilde p(\btheta \, | \, \y) \propto \frac{p(\y, \x, \btheta)}{\tilde p_G(\x \, | \, \btheta, \y)} \Big\rvert_{\x = \bm{\mu}^\star(\btheta)}  \label{eq:hypermarginal}
\end{equation}
which can be marginalised to get $\tilde p(\theta_j \, | \, \y)$
2) In both \eqref{eq:inla1} and \eqref{eq:inla2} we want to integrate w.r.t. \eqref{eq:hypermarginal}, so choose integration points and weights $\{ \btheta^{(k)}, \Delta^{(k)} \}$
  * For low $m$ INLA uses a grid-strategy which I illustrate in the next slide
  * For larger $m$ this becomes too expensive and a CCD design is used

#

```{r message=FALSE, echo=FALSE, fig.cap=paste("An illustration of the INLA grid method for selecting integration points using a toy Gaussian distribution for $\\btheta$. Start at the mode and work outwards along the eigenvectors until the density drops sufficiently low. \\label{fig:grid}")}
knitr::include_graphics("depends/inla-grid.pdf")
```

# Recap on Integrated Nested Laplace Approximation IV

3) Choose approximation for $\tilde p(x_i \, | \, \btheta, \y)$
  * Simplest version [@rue2007approximate] is to marginalise the $p_G(\x \, | \, \btheta, \y)$
\begin{equation}
\tilde p(x_i \, | \, \btheta, \y) = \mathcal{N}(x_i \, | \, \mu^\star_i(\btheta), 1 / q^\star_i(\btheta))
\end{equation}
  * The above is referred to as "Gaussian" approximation, and confusingly there are two more complex ones called "simplified Laplace" and "Laplace"
  * You can pick which one in `R-INLA` using the `method` option
4) Finally use quadrature to get 
\begin{equation}
  \tilde p(x_i \, | \, \y) = 
  \sum_{k = 1}^K \tilde p(x_i \, | \, \btheta^{(k)}, \y) \times \tilde p(\btheta^{(k)} \, | \, \y) \times \Delta^{(k)}
\end{equation}

# Template Model Builder I

* R package which implements the Laplace approximation for latent variable models using AD (via `CppAD`)
  * For more about AD see e.g. @griewank2008evaluating
  * Useful for getting the mode, Hessian
* Write an objective function $f(\x, \btheta)$ in C++ ("user template")
    * We select $f(\x, \btheta) = - \log p(\y \, | \, \x, \btheta) p(\x \, | \, \btheta) p(\btheta)$

# Template Model Builder II

```{cpp eval=FALSE}
#include <TMB.hpp>

template <class Type>
Type objective_function<Type>::operator()() {
  // Define data e.g.
  DATA_VECTOR(y);
  // Define parameters e.g.
  PARAMETER(mu);
  // Calculate negative log-likelihood e.g.
  nll = Type(0.0);
  nll -= dnorm(y, mu, 1, true).sum()
  return(nll);
}
```

# Template Model Builder III

* Performs the Laplace approximation $L_f(\btheta) \approx L^\star_f(\btheta)$ and use R to optimise this with respect to $\btheta$ to give $\hat \btheta$ (the central point in Figure \ref{fig:grid})
  * This is done by specifying the `random` argument to be the parameters that you want to integrate out with a Laplace approximation (the latent field)
* MAP estimate of $\x$ conditional on $\hat \btheta$
* Standard errors calculated using the $\delta$-method (a Gaussian assumption)

# Adaptive Gaussian Hermite Quadrature

* Recent work by Alex Stringer and coauthors on AGHQ
  * `aghq` R package and vignette [@stringer2021implementing]
  * Theory paper [@bilodeau2021stochastic]
* Gauss-Hermite quadrature is a way of picking nodes and weights, and is based on the theory of polynomial interpolation
* The adaptive part means that it uses the location (mode) and curvature (Hessian) of the target (posterior)
* Use $k$ quadrature points
  * If $k$ is odd then they include the mode
  * If $k = 1$ then it's a Laplace approximation
  * In the vignette $k = 3$ (for each dimension, so $3^m$ total) is chosen quite often

# Epil example I

* Epilepsy example from Section 5.2. of @rue2009approximate (previously from BUGS):
  * Patients $i = 1, \ldots, 59$ each either assigned treatment $\texttt{Trt}_i = 1$ or placebo $\texttt{Trt}_i = 0$ to help with seizures
  * Visits to clinics $j = 1, \ldots, 4$ times with $y_{ij}$ the number of seizures of the $i$th person in the two weeks proceeding their $j$th visit to the clinic
  * Covariates age $\texttt{Age}_i$, baseline seizure counts $\texttt{Base}_i$ and an indicator for the final clinic visit $\texttt{V}_4$

# Epil example II

This is what the model looks like (it's a Poisson GLMM):
\begin{align*}
  y_{ij}          &\sim \text{Poisson}(\lambda_{ij}), \\
  \lambda_{ij}    &= e^{\eta_{ij}}, \\
  \eta_{ij}       &= \beta_0 + \beta_{\texttt{Base}} \log(\texttt{Baseline}_j / 4) + \beta_{\texttt{Trt}} \texttt{Trt}_i +
                     \beta_{\texttt{Trt} \times \texttt{Base}} \texttt{Trt}_i \times \log(\texttt{Baseline}_j / 4) \\ 
                  &+ \beta_{\texttt{Age}} \log(\texttt{Age}_i) + \beta_{\texttt{V}_4} {\texttt{V}_4}_j +
                     \epsilon_i + \nu_{ij}, \quad i=1:59, \quad j=1:4, \\
  \beta           &\sim \mathcal{N}(0, 100^2), \quad \forall \beta, \\
  \epsilon_i      &\sim \mathcal{N}(0, 1/\tau_\epsilon), \\
  \nu_{ij}        &\sim \mathcal{N}(0, 1/\tau_\nu), \\
  \tau_\epsilon   &\sim \Gamma(0.001, 0.001), \\
  \tau_\nu        &\sim \Gamma(0.001, 0.001).
\end{align*}

# Epil example III

* `aghq` package interfaces really easily with `TMB`!
* This is the code I used to fit the model with `TMB`

```{r eval=FALSE}
obj <- MakeADFun(
  data = dat,
  parameters = param,
  # These are the ones integrated out with a Laplace approximation
  random = c("epsilon", "nu"),
  DLL = "epil"
)
```

# Epil example IV

* Then to fit it with `aghq` it's only a very small modification

```{r eval=FALSE}
fit <- aghq::marginal_laplace_tmb(
  obj, 
  k = 3, 
  startingvalue = c(param$beta, param$l_tau_epsilon, param$l_tau_nu)
)
```

#

```{r echo=FALSE}
results <- readRDS("comparison-results.rds")

beta_i <- function(i) { c(paste0("beta_", i), paste0("sd(beta_", i, ")")) }
rownames(results) <- c(sapply(0:5, beta_i))

results %>% 
  round(digits = 3) %>%
  kableExtra::kable(format = "latex", booktabs = TRUE, escape = TRUE, align = "c") %>%
  kable_styling(latex_options = "scale_down")
```

# Plan

* Test `aghq` for toy Naomi example
  * Do as above with the Epil example, testing versus a long MCMC run
  * Sometimes you have to look pretty hard for a node (element of the latent field) where there are differences. In the INLA paper they do this by computing a SKLD and ordering by maximum difference. Probably good to do here as well
* Extend `aghq` to replicate INLA functionality by adding the more complex versions of $\tilde p(x_i \, | \, \btheta, \y)$ then test that wih Naomi
  * Håvard Rue philosophy: "do one thing and do it well"
  * `R-INLA` implementation of INLA based on sparsity of $\bm{Q}(\btheta)$ that doesn't hold up for extended LGMs
  * @wood2020simplified on how to still do it
* Try the INLA without `R-INLA` on other almost LGMs and see how far it can be pushed

# References {.allowframebreaks}