### 2022-09-07 src/prev-anc-art_process-results/prev-anc-art.Rmd
prev-anc-art html up

---
title: "Prevalence, ANC, ART model"
author:
- name: Adam Howes
output:
  html_document:
    toc: yes
    toc_float: true
    toc_collapsed: true
    df_print: paged
    code_folding: show
    theme: lumen
---

```{r setup, class.source = 'fold-hide'}
knitr::opts_chunk$set(
  cache = TRUE,
  autodep = TRUE,
  cache.lazy = FALSE,
  cache.comments = FALSE
)
options(scipen = 999)
cbpalette <- multi.utils::cbpalette()
library(tidyverse)
rstan::rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

source("functions.R")
```

Create true values dataframe:

```{r}
true_values <- data.frame(
  parameter = c(
    "beta_prev",
    "log_sigma_phi_prev",
    "beta_anc",
    "log_sigma_b_anc",
    "beta_art",
    "log_sigma_phi_art"
  ),
  true_value = c(
    -2.4,
    log(sqrt(1 / 4)),
    -0.2,
    log(sqrt(1/ 100)),
    0.7,
    log(0.35)
  )
)
```

Read in results created in reports:

* [Model 0](https://github.com/athowes/elgm-inf/tree/master/src/prev-anc-art_model0)
* [Model 1](https://github.com/athowes/elgm-inf/tree/master/src/prev-anc-art_model1)
* [Model 2](https://github.com/athowes/elgm-inf/tree/master/src/prev-anc-art_model2)
* [Model 3](https://github.com/athowes/elgm-inf/tree/master/src/prev-anc-art_model3)
* [Model 4](https://github.com/athowes/elgm-inf/tree/master/src/prev-anc-art_model4)

```{r}
# model0 <- readRDS("depends/results_model0.rds")
model1 <- readRDS("depends/results_model1.rds")
model2 <- readRDS("depends/results_model2.rds")
model3 <- readRDS("depends/results_model3.rds")
model4 <- readRDS("depends/results_model4.rds")
```

## Boxplots

```{r cache=FALSE}
# draw_boxplots(model0) + labs(title = "Model 0")
# draw_boxplots(model1) + labs(title = "Model 1")
# draw_boxplots(model2) + labs(title = "Model 2")
# draw_boxplots(model3) + labs(title = "Model 3")
# draw_boxplots(model4) + labs(title = "Model 4")
```

## Scatterplots

```{r fig.height=10}
# draw_scatterplots(model0) + labs(title = "Model 0")
draw_scatterplots(model1) + labs(title = "Model 1")
draw_scatterplots(model2) + labs(title = "Model 2")
draw_scatterplots(model3) + labs(title = "Model 3")
draw_scatterplots(model4) + labs(title = "Model 4")
```

## Kolmogorov KS test plots

Take samples from all distributions, then compute maximum ECDF difference $D$ (two-sample Kolmogorov–Smirnov test).
We assume that `tmbstan` represents the gold-standard, and compare performance of `aghq` and `TMB` via their similarity to this gold-standard.
Higher values of $D$ correspond to distributions which are more different.

```{r fig.height=8}
# draw_ksplots_D(model0) + labs(title = "Model 0")
draw_ksplots_D(model1) + labs(title = "Model 1")
draw_ksplots_D(model2) + labs(title = "Model 2")
draw_ksplots_D(model3) + labs(title = "Model 3")
draw_ksplots_D(model4) + labs(title = "Model 4")
```

We could also assess $l$, the location of $D$.
Determining if there are patterns in the location of the greatest ECDF difference could present us with useful insights.

```{r fig.height=8}
# draw_ksplots_D(model0) + labs(title = "Model 0")
draw_ksplots_l(model1) + labs(title = "Model 1")
draw_ksplots_l(model2) + labs(title = "Model 2")
draw_ksplots_l(model3) + labs(title = "Model 3")
draw_ksplots_l(model4) + labs(title = "Model 4")
```

## MCMC diagnostics

When using Markov chain Monte Carlo (MCMC) methods, as we have for `tmbstan`, it's important to assess for convergence.

### Traceplots

One way to do this is via traceplots which visualise the chains over the number of iterations specified.

```{r}
# map(model0, "mcmc_traceplots")
map(model1, "mcmc_traceplots")
map(model2, "mcmc_traceplots")
map(model3, "mcmc_traceplots")
map(model4, "mcmc_traceplots")
```

### R hat

The $\hat R$ statistic ("R hat") can also be used.
A value of $\hat R < 1.1$ is typically sufficient.

```{r}
# draw_rhatplot(model0) + labs(title = "Model 0")
draw_rhatplot(model1) + labs(title = "Model 1")
draw_rhatplot(model2) + labs(title = "Model 2")
draw_rhatplot(model3) + labs(title = "Model 3")
draw_rhatplot(model4) + labs(title = "Model 4")
```