### 2022-10-06 src/explore_aghq/walkthrough.Rmd
silent = TRUE to make this more readable

---
title: "Code walkthrough"
author:
- name: Adam Howes
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Background** The `aghq` package implements adaptive Gaussian Hermite quadrature. It is particularly compatible with the `TMB` package for Laplace approximations.

  **Task** In preparation for implementing new methods into `aghq`, in this notebook we walkthrough various `aghq` functions. This will improve understanding of (1) the method, and (2) how to manipulate `TMB` objects in R.
---

We use the epilepsy example, explained in more detail in the [`epil`](https://athowes.github.io/elgm-inf/epil.html) notebook, as example.
Start by loading and preparing the data, as well as compling the `TMB` template:

```{r}
data(Epil, package = "INLA")

center <- function(x) (x - mean(x))

Epil <- Epil %>%
  mutate(CTrt    = center(Trt),
         ClBase4 = center(log(Base/4)),
         CV4     = center(V4),
         ClAge   = center(log(Age)),
         CBT     = center(Trt * log(Base/4)))

N <- 59
J <- 4
K <- 6
X <- model.matrix(formula(~ 1 + CTrt + ClBase4 + CV4 + ClAge + CBT), data = Epil)
y <- Epil$y

make_epsilon_matrix <- function(N, J) {
  t(outer(1:N, 1:(N * J), function(r, c) as.numeric((J*(r - 1) < c) & (c <= J*r))))
}

E <- make_epsilon_matrix(N = 3, J = 2)

dat <- list(N = N, J = J, K = K, X = X, y = y, E = make_epsilon_matrix(N, J))

compile("epil.cpp")
dyn.load(dynlib("epil"))

param <- list(
  beta = rep(0, K),
  epsilon = rep(0, N),
  nu = rep(0, N * J),
  l_tau_epsilon = 0,
  l_tau_nu = 0
)

obj <- MakeADFun(
  data = dat,
  parameters = param,
  random = c("epsilon", "nu"),
  DLL = "epil",
  silent = TRUE
)
```

# `aghq::marginal_laplace_tmb`

We start by calling this function, to show where we're trying to end up:

```{r results='hide'}
fit <- aghq::marginal_laplace_tmb(
  obj,
  k = 2,
  startingvalue = c(param$beta, param$l_tau_epsilon, param$l_tau_nu)
)
```

```{r}
summary(fit)
```

And now let's go through line by line.
These are all of the inputs to the function:

```{r}
ff <- obj
k <- 2
startingvalue <- c(param$beta, param$l_tau_epsilon, param$l_tau_nu)
transformation <- aghq::default_transformation()
optresults <- NULL
basegrid <- NULL
control <- aghq::default_control_tmb()
```

```{r}
#' My guess is that this function just makes sure that the control input is valid
validate_control(control, type = "tmb")

#' And similarly, this probably makes sure the transformation input is valid
validate_transformation(transformation)

transformation <- make_transformation(transformation)
transformation$totheta
transformation$fromtheta
transformation$jacobian

#' Get names for the theta parameters from the TMB function template
#' Note that make.unique will make the names unique by apending sequence numbers to duplicates
thetanames <- NULL
if (exists('par', ff)) thetanames <- make.unique(names(ff$par), sep = "")
thetanames

#' Hessian
if (control$numhessian) {
  ff$he <- function(theta) numDeriv::jacobian(ff$gr, theta, method = "Richardson")
}
```

```{r results='hide'}
#' Now we're going to do the AGHQ
quad <- aghq(
  ff = ff,
  k = k,
  transformation = transformation,
  startingvalue = startingvalue,
  optresults = optresults,
  basegrid = basegrid,
  control = control
)
```

```{r}
#' There is an option whereby quad will just return the normalising constant, and if so then tmb_marginal_laplace can also be used
#' just to return the normalising constant too. Note on this, why use tmb_marginal_laplace in this situation rather than aghq? Since
#' as far as I understand we're not doing the marginal laplace approximation if control$onlynormconst is TRUE
if (control$onlynormconst) return(quad)

distinctthetas <- quad$normalized_posterior$nodesandweights[, grep('theta', colnames(quad$normalized_posterior$nodesandweights))]
distinctthetas
```

Note that there are `r nrow(distinctthetas)` "distinct thetas" corresponding to $k$, here `r k`, raised to the power `r ncol(distinctthetas)`:

```{r}
dim(distinctthetas)
```

```{r}
#' I think this will just be fixing a case where there is only one theta, and somehow a data.frame isn't created
if (!is.data.frame(distinctthetas)) distinctthetas <- data.frame(theta1 = distinctthetas)

modesandhessians <- distinctthetas

#' Replacing the theta, theta1, theta2, ... names with the names from TMB
if (is.null(thetanames)) {
  thetanames <- colnames(distinctthetas)
} else {
  colnames(modesandhessians) <- thetanames
  colnames(quad$normalized_posterior$nodesandweights)[grep('theta',colnames(quad$normalized_posterior$nodesandweights))] <- thetanames
}

modesandhessians$mode <- vector(mode = "list", length = nrow(distinctthetas))
modesandhessians$H <- vector(mode = "list", length = nrow(distinctthetas))

#' There is a loop coming up, which we will step through for i = 1, then let a loop do for i = 2, ...
i <- 1

#' Get the theta (currently modesandhessians, despite being called modesandhessians, only contains the theta, not any mode or hessian)
theta <- as.numeric(modesandhessians[i, thetanames])
theta

#' Set up the mode and hessian of the random effects. This happens when you run the TMB objective with a particular theta
ff$fn(theta)

#' Now pull the mode and hessian. Have to be careful about scoping

#' So this is the mode, for both random and fixed effects
mm <- ff$env$last.par

#' These are the indices of the random effects
ff$env$random

#' We can confirm that for the fixed effects, the "mode" is just the input values theta
#' (To get the indicies for the fixed effects, we just take the complement of the random effects)
mm[-ff$env$random]
theta

#' Set the mode using the relevant indicies of mm
#' "mode" is a list column
modesandhessians[i, "mode"] <- list(list(mm[ff$env$random]))
modesandhessians[i, ]

#' This should be the Hessian evaluated at the mode
#' The argument random = TRUE set it so that...
H <- ff$env$spHess(mm, random = TRUE)

#' Alex has the following note on duplicating H here:
#' "Somehow, TMB puts all evaluations of spHess in the same memory location"
#' Don't totally understand this
H <- rlang::duplicate(H)

modesandhessians[i, "H"] <- list(list(H))

for (i in 2:nrow(distinctthetas)) {
  theta <- as.numeric(modesandhessians[i,thetanames])
  ff$fn(theta)
  mm <- ff$env$last.par
  modesandhessians[i,'mode'] <- list(list(mm[ff$env$random]))
  H <- ff$env$spHess(mm,random = TRUE)
  H <- rlang::duplicate(H)
  modesandhessians[i,'H'] <- list(list(H))
}

quad$modesandhessians <- modesandhessians

class(quad) <- c("marginallaplace", "aghq")
```