### 2022-10-08 src/explore_aghq/walkthrough.Rmd
aghq::sample_margial step-by-step

---
title: "Code walkthrough"
author:
- name: Adam Howes
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Background** The `aghq` package implements adaptive Gaussian Hermite quadrature. It is particularly compatible with the `TMB` package for Laplace approximations.

  **Task** In preparation for implementing new methods into `aghq`, in this notebook we walkthrough various `aghq` functions. This will improve understanding of (1) the method, and (2) how to manipulate `TMB` objects in R.
---

We use the epilepsy example, explained in more detail in the [`epil`](https://athowes.github.io/elgm-inf/epil.html) notebook, as example.
Start by loading and preparing the data, as well as compling the `TMB` template:

```{r}
data(Epil, package = "INLA")

center <- function(x) (x - mean(x))

Epil <- Epil %>%
  mutate(CTrt    = center(Trt),
         ClBase4 = center(log(Base/4)),
         CV4     = center(V4),
         ClAge   = center(log(Age)),
         CBT     = center(Trt * log(Base/4)))

N <- 59
J <- 4
K <- 6
X <- model.matrix(formula(~ 1 + CTrt + ClBase4 + CV4 + ClAge + CBT), data = Epil)
y <- Epil$y

make_epsilon_matrix <- function(N, J) {
  t(outer(1:N, 1:(N * J), function(r, c) as.numeric((J*(r - 1) < c) & (c <= J*r))))
}

E <- make_epsilon_matrix(N = 3, J = 2)

dat <- list(N = N, J = J, K = K, X = X, y = y, E = make_epsilon_matrix(N, J))

compile("epil.cpp")
dyn.load(dynlib("epil"))

param <- list(
  beta = rep(0, K),
  epsilon = rep(0, N),
  nu = rep(0, N * J),
  l_tau_epsilon = 0,
  l_tau_nu = 0
)

obj <- MakeADFun(
  data = dat,
  parameters = param,
  random = c("epsilon", "nu"),
  DLL = "epil",
  silent = TRUE
)
```

# `aghq::marginal_laplace_tmb`

We start by calling this function, to show where we're trying to end up:

```{r results='hide'}
fit <- aghq::marginal_laplace_tmb(
  obj,
  k = 2,
  startingvalue = c(param$beta, param$l_tau_epsilon, param$l_tau_nu)
)
```

```{r}
summary(fit)
```

And now let's go through line by line.
These are all of the inputs to the function:

```{r}
ff <- obj
k <- 2
startingvalue <- c(param$beta, param$l_tau_epsilon, param$l_tau_nu)
transformation <- aghq::default_transformation()
optresults <- NULL
basegrid <- NULL
control <- aghq::default_control_tmb()
```

```{r}
#' My guess is that this function just makes sure that the control input is valid
validate_control(control, type = "tmb")

#' And similarly, this probably makes sure the transformation input is valid
validate_transformation(transformation)

transformation <- make_transformation(transformation)
transformation$totheta
transformation$fromtheta
transformation$jacobian

#' Get names for the theta parameters from the TMB function template
#' Note that make.unique will make the names unique by apending sequence numbers to duplicates
thetanames <- NULL
if (exists('par', ff)) thetanames <- make.unique(names(ff$par), sep = "")
thetanames

#' Hessian
if (control$numhessian) {
  ff$he <- function(theta) numDeriv::jacobian(ff$gr, theta, method = "Richardson")
}
```

```{r results='hide'}
#' Now we're going to do the AGHQ
quad <- aghq(
  ff = ff,
  k = k,
  transformation = transformation,
  startingvalue = startingvalue,
  optresults = optresults,
  basegrid = basegrid,
  control = control
)
```

```{r}
#' There is an option whereby quad will just return the normalising constant, and if so then tmb_marginal_laplace can also be used
#' just to return the normalising constant too. Note on this, why use tmb_marginal_laplace in this situation rather than aghq? Since
#' as far as I understand we're not doing the marginal laplace approximation if control$onlynormconst is TRUE
if (control$onlynormconst) return(quad)

distinctthetas <- quad$normalized_posterior$nodesandweights[, grep('theta', colnames(quad$normalized_posterior$nodesandweights))]
distinctthetas
```

Note that there are `r nrow(distinctthetas)` "distinct thetas" corresponding to $k$, here `r k`, raised to the power `r ncol(distinctthetas)`:

```{r}
dim(distinctthetas)
```

```{r}
#' I think this will just be fixing a case where there is only one theta, and somehow a data.frame isn't created
if (!is.data.frame(distinctthetas)) distinctthetas <- data.frame(theta1 = distinctthetas)

modesandhessians <- distinctthetas

#' Replacing the theta, theta1, theta2, ... names with the names from TMB
if (is.null(thetanames)) {
  thetanames <- colnames(distinctthetas)
} else {
  colnames(modesandhessians) <- thetanames
  colnames(quad$normalized_posterior$nodesandweights)[grep('theta', colnames(quad$normalized_posterior$nodesandweights))] <- thetanames
}

modesandhessians$mode <- vector(mode = "list", length = nrow(distinctthetas))
modesandhessians$H <- vector(mode = "list", length = nrow(distinctthetas))

#' There is a loop coming up, which we will step through for i = 1, then let a loop do for i = 2, ...
i <- 1

#' Get the theta (currently modesandhessians, despite being called modesandhessians, only contains the theta, not any mode or hessian)
theta <- as.numeric(modesandhessians[i, thetanames])
theta

#' Set up the mode and hessian of the random effects. This happens when you run the TMB objective with a particular theta
ff$fn(theta)

#' Now pull the mode and hessian. Have to be careful about scoping

#' So this is the mode, for both random and fixed effects
mm <- ff$env$last.par

#' These are the indices of the random effects
ff$env$random

#' We can confirm that for the fixed effects, the "mode" is just the input values theta
#' (To get the indicies for the fixed effects, we just take the complement of the random effects)
mm[-ff$env$random]
theta

#' Set the mode using the relevant indicies of mm
#' "mode" is a list column
modesandhessians[i, "mode"] <- list(list(mm[ff$env$random]))
modesandhessians[i, ]

#' This should be the Hessian evaluated at the mode
#' The argument random = TRUE set it so that...
H <- ff$env$spHess(mm, random = TRUE)

#' Alex has the following note on duplicating H here:
#' "Somehow, TMB puts all evaluations of spHess in the same memory location"
#' Don't totally understand this
H <- rlang::duplicate(H)

modesandhessians[i, "H"] <- list(list(H))

for (i in 2:nrow(distinctthetas)) {
  theta <- as.numeric(modesandhessians[i,thetanames])
  ff$fn(theta)
  mm <- ff$env$last.par
  modesandhessians[i,'mode'] <- list(list(mm[ff$env$random]))
  H <- ff$env$spHess(mm,random = TRUE)
  H <- rlang::duplicate(H)
  modesandhessians[i,'H'] <- list(list(H))
}

quad$modesandhessians <- modesandhessians

class(quad) <- c("marginallaplace", "aghq")

```

# `aghq::sample_marginal.marginallaplace`

What we'd like to do is to sample from
$$
\begin{align}
\tilde p(x_i \, | \, y) &= \sum_z \tilde p_\text{G}(x_i \, | \, \theta(z), y) \tilde p_\text{LA}(\theta(z) \, | \, y) w_k(z), \\
&= \sum_z \tilde p_\text{G}(x_i \, | \, \theta(z), y) \lambda(z)
\end{align}
$$
where $\lambda(z) \in [0, 1]$ are weights such that $\sum_z \lambda(z) = 1$.
We can do this by

1. Sample $z_j \sim \text{Multinomial}(1, \lambda(z_1), \ldots, \lambda(z_J))$
2. Given $z_j$ sample from $\tilde p_\text{G}(x \, | \, \theta(z_j), y)$
3. Take the $i$th entry of $\tilde p_\text{G}(x \, | \, \theta(z_j), y)$ as the sample from $\tilde p_\text{G}(x_i \, | \, \theta(z_j), y)$

Now that we have `quad`, it should be compatible with `sample_marginal`:

```{r}
#' The number of samples to get
M <- 10

transformation <- quad$transformation
interpolation <- "auto"

#' Is this getting mc.cores, and if there isn't a value just setting it to 1 (integer)
numcores <- getOption("mc.cores", 1L)

#' Length of the random effects
d <- dim(quad$modesandhessians$H[[1]])[1]

simlist <- quad$modesandhessians

#' Avoid use of parallel computing on Windows
if (.Platform$OS.type == 'windows') numcores <- 1
```

The next step is to compute the Cholesky decomposition `L` of the Hessian `H` such that
$$
H = LL^\top
$$
The reason we want the Cholesky is that we will be using it to same from the joint Gasusian approximation.
The function `chol` from `base` will compute it for us.
For example, for the first element:

```{r}
h <- simlist$H[[1]]
h <- Matrix::forceSymmetric(h)
L <- chol(h, perm = FALSE)
L
```

Doing this for the list using `lapply`:

```{r}
simlist$L <- lapply(simlist$H,function(h) chol(Matrix::forceSymmetric(h), perm = FALSE))
```

There is another version, using the arguments `perm` and `LDL` (which I don't understand) that may be quicker but ultimately will achieve the same thing:

```{r}
# simlist$L <- lapply(simlist$H, function(h) Matrix::Cholesky(as(Matrix::forceSymmetric(h), "sparseMatrix"), perm = TRUE, LDL = FALSE))
```

Now let's compute the $\lambda(z)$ weights.

```{r}
simlist$lambda <- exp(quad$normalized_posterior$nodesandweights$logpost_normalized) * quad$normalized_posterior$nodesandweights$weights
head(simlist$lambda)
```

We can check that they do indeed sum to one:

```{r}
sum(simlist$lambda)
```

Now let's sample from the multinomial:

```{r}
j <- apply(stats::rmultinom(M, 1, simlist$lambda), 2, function(x) which(x == 1))
length(j) #' Should be M
j

#' Values are number of samples to draw for each j
#' In this case they're likely to be all 1, since we're not doing many samples
tt <- table(j)
```

Now we construct a big Gaussian mixture matrix.
The first step is to create a matrix with `d` columns, one for each random effect, and `M` rows, one for each sample, where each value is a standard normal $\mathcal{N}(0, 1)$:

```{r}
A <- matrix(stats::rnorm(M * d), nrow = M)

#' Now make a list where each element is a row of A, where the index is determined by j
A_split <- split(A, j)

#' Big Gaussian mixture matrix

#' For each element of the list, create a matrix with d rows (and 1 column? always?)
Z <- lapply(A_split, matrix , nrow = d)
str(Z)
```

In the code there is now a big call to `mapply`, which we are going to break down step-by-step.
First set the inital values that `function(.x, .y)` will iterate over:

```{r}
.x <- Z[[1]]
str(.x)

.y <- names(Z)[[1]]
str(.y)
```

Now solve the equation `L %*% .z = .x` for `.z`

```{r}
.z <- solve(simlist$L[[as.numeric(.y)]], .x)
str(.z)
```

Now add the mode (where the additional manipulation is so that if there are multiple samples from the same `j` then there will be the corresonding number of columns):

```{r}
.z <- .z + do.call(cbind, rep(list(simlist$mode[[as.numeric(.y)]]), ncol(.x)))
str(.z)
```

Do this using a `mapply` call:

```{r}
samps <- mapply(
  function(.x, .y) as.numeric(solve(simlist$L[[as.numeric(.y)]], .x)) + do.call(cbind, rep(list(simlist$mode[[as.numeric(.y)]]), ncol(.x))),
  Z,
  names(Z),
  SIMPLIFY = FALSE
)

#' Order them properly
#' My guess is that you'd want them to be ordered according to how j was drawn, rather
#' than being grouped up by the value of lambda
ord <- numeric(length(j))
cumtab <- cumsum(c(0, tt))
cumtab <- cumtab[-length(cumtab)]
cnt <- numeric(length(unique(j)))
names(cnt) <- sort(unique(j))

for (i in 1:length(j)) {
  wc <- which(names(cnt) == j[i])
  cnt[wc] <- cnt[wc] + 1
  ord[i] <- cumtab[wc] + cnt[wc]
}

samps <- Reduce(cbind, samps)
samps <- samps[, ord]
samps
```

```{r}
#' This is now the length of the fixed effects
d <- length(quad$optresults$mode)

#' Just so that we can get the names of the theta (fixed effects)
thetanames <- colnames(quad$normalized_posterior$nodesandweights)[1:d]
theta <- simlist[j, thetanames]

#' In one dimension, R's indexing is not type consistent
if (!is.matrix(samps)) {
  samps <- rbind(samps)
  rownames(samps) <- NULL
}

if (!inherits(theta, "data.frame")) theta <- data.frame(theta1 = theta)

out <- list(
  samps = samps,
  theta = theta
)

out$theta

# Add the marginals for theta, with possible transformation
class(quad) <- "aghq" #' This is preventing sample_marginal.marginallaplace being used
out$thetasamples <- sample_marginal(quad, M, transformation, interpolation)
str(out)
```

# `aghq::sample_marginal.aghq`

Let's have a look at what that call to `sample_marginal.aghq` is doing:

```{r}
class(quad)

out <- list()

#' If the quadrature object somehow doesn't contain any information about the 
if (is.null(quad$marginals)) return(out)

str(quad$marginals)
```

As usual, we're going to start by doing the computation for the first element of `theta` explicitly:

```{r}
i <- 1
```

We already have information about each marginal, contained in:

```{r}
quad$marginals[[i]]
```

So we can use the function `aghq::compute_quantiles` to, my guess, use the inverse CDF method to sample from the marginal, given input of some $\mathcal{U}(0, 1)$ draws:

```{r}
unif <- stats::runif(M)
unif

compute_quantiles(quad$marginals[[i]], unif, interpolation = interpolation, transformation = transformation)
```

```{r}
for (i in 1:length(quad$marginals)) {
  out[[i]] <- unname(compute_quantiles(quad$marginals[[i]], stats::runif(M), interpolation = interpolation, transformation = transformation))
}

#' The function doesn't name these yet, but in future it could (and also return a data.frame) via
d <- length(quad$optresults$mode)
thetanames <- colnames(quad$normalized_posterior$nodesandweights)[1:d]
names(out) <- thetanames
as.data.frame(out)
```

# `aghq::compute_quantiles.aghq`

Does the `compute_quantiles` function work as we expect it to?

```{r}
class(quad$marginals[[i]])
```

The PDF and CDF are computed by `compute_pdf_and_cdf` (as the name suggests, and another layer deep that we can look).
Code comments note that transformations are not applied at this point, and that it's better to transform the quantiles later:

```{r}
q <- stats::runif(M)

pdfandcdf <- compute_pdf_and_cdf(quad$marginals[[i]], interpolation = interpolation)

ggplot(pdf_and_cdf, aes(x = theta, y = cdf)) +
  geom_line() +
  theme_minimal()

out <- numeric(length(q))
increasing <- TRUE

validate_transformation(transformation)
transformation <- make_transformation(transformation)
```

For each sample, take the maximum `theta` which still has CDF value less than the quantile desired:

```{r}
for (i in 1:length(q)) out[i] <- pdfandcdf$theta[max(which(pdfandcdf$cdf < q[i]))]

#' Have to check if it's increasing or decreasing so can reverse order if necessary
increasing <- transformation$fromtheta(min(out)) <= transformation$fromtheta(max(out))
for (i in 1:length(out)) out[i] <- transformation$fromtheta(out[i])

if (!increasing) out <- rev(out)
names(out) <- paste0(as.character(100 * q), "%")
out
```

# `compute_pdf_and_cdf`

Going to leave this as a to-do.