### 2022-10-12 src/docs_01-04-20-mini/01-04-20-mini.Rmd
Upload old mini project presentation

---
title: "Mini-project: Integrated nested Laplace Approximation with Automatic Differentiation"
author: "Adam Howes"
date: "April 2020"
output:
  beamer_presentation:
    fig_height: 2.5
    fig_width: 6
    highlight: haddock
    includes:
      in_header: preamble.tex
    latex_engine: pdflatex
  slidy_presentation:
    highlight: haddock
institute: Imperial College London
bibliography: citations.bib
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
```

## Motivation

* `R-INLA` [@martins2013bayesian] only works for the particular models which have been implemented
* Alternative implementation based on automatic differentation (AD) would allow INLA to be used for a \textcolor{hilit}{broader class} of models
* For example, the HIV inference group at Imperial is working on a model just outside `R-INLA`'s capacity

# What is INLA, why do we want to use it, and why can't we currently?

## Three-stage model

* Want to do Bayesian inference in spatiotemporal statistics
* \textcolor{hilit}{Three-stage model} covers most of the models used
\begin{alignat*}{2}
  &\text{(Observations)}     &   \y &\sim p(\y \, | \, \x), \\
  &\text{(Latent field)}     &   \x &\sim p(\x \, | \, \btheta), \\
  &\text{(Hyperparameters)}  &   \qquad \btheta &\sim p(\btheta),
\end{alignat*}
where $\y = (y_1, \ldots, y_n)$, $\x = (x_1, \ldots, x_n)$, $\btheta = (\theta_1, \ldots, \theta_m)$
* Interested in learning \textcolor{hilit}{both} $(\btheta, \x)$ from data $\y$

## Have you tried MCMC?

* Markov chain Monte Carlo is slow for high dimensional correlated parameter spaces
* We have both of these problems:
  * If $\x$ represents spatiotemporal location then $\text{dim}(\x) = n$ will be very large 
  * Tobler's first law of geography "everything is related to everything else, but near things are more related than distant things" $\implies \x$ has lots of correlation structure

## Approximate Bayesian inference

* In applied statistics (at least in health and social science) we fit misspecified models to biased and incomplete data
* Is inferential exactness (as $n_\text{sim} \to \infty$ for chain of length $n_\text{sim}$) the scientific bottleneck?
* If not $\implies$ shouldn't be afraid of approximate methods
    * Approximate Bayesian computation (ABC)
    * Variational Bayes
    * \textcolor{hilit}{Integrated nested Laplace approximation} (INLA)

## Integrated nested Laplace approximation (I)

* See @rue2009approximate or @blangiardo2015spatial
* Approximate Bayesian inference for \textcolor{hilit}{latent Gaussian models} (LGMs), which are three-stage models with middle layer
\begin{equation*}
  \text{(Latent field)} \qquad  p(\x \, | \, \btheta) = \mathcal{N}(\x  \, | \, \bm{\mu}(\btheta), \bm{Q}(\btheta)^{-1}).
\end{equation*}
* Takes advantage of sparsity properties of $\bm{Q}(\btheta)$, i.e. if $\x$ is a Gaussian Markov random field (GMRF)
* Gives approximate \textcolor{hilit}{posterior marginals} $\{\tilde p(x_i \, | \, \y)\}_{i = 1}^n$ and $\{\tilde p(\theta_j \, | \, \y)\}_{j = 1}^m$

## Integrated nested Laplace approximation (II)

1) First Laplace approximate hyperparameter posterior
\begin{equation}
\tilde p(\btheta \, | \, \y) \propto \frac{p(\y \, | \, \x, \btheta) p(\x \, | \, \btheta) p(\btheta)}{\tilde p_G(\x \, | \, \btheta, \y)} \Big\rvert_{\x = \bm{\mu}^\star(\btheta)}  \label{eq:hypermarginal}
\end{equation}
which can be marginalised to get $\tilde p(\theta_j \, | \, \y)$
2) Choose integration points and weights $\{ \btheta^{(k)}, \Delta^{(k)} \}$ to integrate w.r.t. \eqref{eq:hypermarginal}
3) Choose approximation for $\tilde p(x_i \, | \, \btheta, \y)$ (simplest version: Gaussian)
4) Finally use quadrature to get 
\begin{equation}
  \tilde p(x_i \, | \, \y) = 
  \sum_{k = 1}^K \tilde p(x_i \, | \, \btheta^{(k)}, \y) \times \tilde p(\btheta^{(k)} \, | \, \y) \times \Delta^{(k)}
\end{equation}

## Naomi, evidence synthesis for HIV

:::::::::::::: {.columns}

::: {.column width=.65}

* Combine HIV prevalence $\rho_i$ and ART coverage $\alpha_i$ models together
* Model is close to, but not, a LGM
* Small non-linearities e.g. multiplying two latent Gaussian fields 
\begin{equation*}
A_i \sim \text{Bin}(\rho_i \alpha_i, N_i),
\end{equation*}
where $A_i$ be the number observed on ART and $N_i$ the population
* Need something more flexible than `R-INLA`

:::

::: {.column width=.35}

```{r, echo=FALSE, fig.cap="Supermodel", out.width = '65%'}
knitr::include_graphics("naomi_hex.png")
```

:::

::::::::::::::

# What do we do currently instead?

## Template Model Builder (I)

* Currently we use `TMB` [@tmb]
* R package which implements the Laplace approximation for latent variable models using AD (via `CppAD`)
  * For more about AD see e.g. @griewank2008evaluating
* Write an objective function $f(\x, \btheta)$ in C++ ("user template")
    * We select $f(\x, \btheta) = - \log p(\y \, | \, \x, \btheta) p(\x \, | \, \btheta) p(\btheta)$

## Template Model Builder (II)

```{cpp eval=FALSE}
#include <TMB.hpp>

template <class Type>
Type objective_function<Type>::operator()() {
  // Define data e.g.
  DATA_VECTOR(y);
  // Define parameters e.g.
  PARAMETER(mu);
  // Calculate negative log-likelihood e.g.
  nll = Type(0.0);
  nll -= dnorm(y, mu, 1, true).sum()
  return(nll);
}
```

## Template Model Builder (III)

* Performs the Laplace approximation $L_f(\btheta) \approx L^\star_f(\btheta)$ (step 1 of INLA) -- use R to optimise this with respect to $\btheta$ to give $\hat \btheta$
* MAP estimate of $\x$ conditional on $\btheta = \hat \btheta$ (REML? Empirical Bayes?)
* Standard errors calculated using the $\delta$-method (a Gaussian assumption)

# What do you want to do in the future?

## Aims

* Compare accuracy of `TMB` to `R-INLA`
* Implement the INLA method using AD via `TMB`
* Apply new method to models with different degrees of non-linearity
    * Small degree: Naomi. 
    * Larger degree: ODE models e.g. SIR or other compartmental models

# Thanks! Questions / comments / corrections?

## References {.allowframebreaks}