### 2022-10-17 src/example_naomi/naomi.Rmd
Upload start on Naomi vignette

---
title: "Naomi model workflow"
author:
- name: Adam Howes
output:
  html_document:
    toc: yes
    toc_float: true
    toc_collapsed: true
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
    **Background** A vignette for the Naomi model is uploaded to the package website.

    **Task** We run through the vignette manually to (1) develop a better understanding of the model, (2) get experience working with the Malawi data, which we will likely use for our case-study as it is publicly available.
---

```{r}
library(naomi)
library(tidyverse)
library(sf)
```

# Naomi vignette

Following vignette [here](https://mrc-ide.github.io/naomi/articles/model-workflow.html).

## Load data inputs

Area hierarchy and boundaries:

```{r}
area_merged <- read_sf(system.file("extdata/demo_areas.geojson", package = "naomi"))

area_merged %>%
  filter(area_level == max(area_level)) %>%
  ggplot() +
  geom_sf(aes(geometry = geometry)) +
  theme_minimal()
```

Population data:

```{r}
pop_agesex <- read_csv(system.file("extdata/demo_population_agesex.csv", package = "naomi"))
head(pop_agesex)
```

Survey data:

```{r}
survey_hiv_indicators <- read_csv(system.file("extdata/demo_survey_hiv_indicators.csv", package = "naomi"))
head(survey_hiv_indicators)
```
ART programmatic data:

```{r}
art_number <- read_csv(system.file("extdata/demo_art_number.csv", package = "naomi"))
head(art_number)
```

ANC programmatic data:

```{r}
anc_testing <- read_csv(system.file("extdata/demo_anc_testing.csv", package = "naomi"))
head(anc_testing)
```

## Choose model areas and time points

```{r}
scope <- "MWI"
level <- 4
calendar_quarter_t1 <- "CY2016Q1"
calendar_quarter_t2 <- "CY2018Q3"
calendar_quarter_t3 <- "CY2019Q4"
```

```{r}
prev_survey_ids  <- c("DEMO2016PHIA", "DEMO2015DHS")
artcov_survey_ids  <- "DEMO2016PHIA"
vls_survey_ids <- NULL
recent_survey_ids <- "DEMO2016PHIA"

artnum_calendar_quarter_t1 <- "CY2016Q1"
artnum_calendar_quarter_t2 <- "CY2018Q3"

anc_clients_year2 <- 2018
anc_clients_year2_num_months <- 9

anc_prevalence_year1 <- 2016
anc_prevalence_year2 <- 2018

anc_art_coverage_year1 <- 2016
anc_art_coverage_year2 <- 2018
```

## Prepare model inputs

```{r eval=FALSE}
naomi_mf <- naomi_model_frame(
  area_merged,
  pop_agesex,
  spec,
  scope = scope,
  level = level,
  calendar_quarter_t1,
  calendar_quarter_t2,
  calendar_quarter_t3
)
```

```{r eval=FALSE}
naomi_data <- select_naomi_data(
  naomi_mf,
  survey_hiv_indicators,
  anc_testing,
  art_number,
  prev_survey_ids,
  artcov_survey_ids,
  recent_survey_ids,
  vls_survey_ids,
  artnum_calendar_quarter_t1,
  artnum_calendar_quarter_t2,
  anc_prevalence_year1,
  anc_prevalence_year2,
  anc_art_coverage_year1,
  anc_art_coverage_year2
)
```

## Fit the TMB model

```{r eval=FALSE}
fit <- fit_tmb(tmb_inputs)
```

```{r eval=FALSE}
outputs <- output_package(fit, naomi_mf)
names(outputs)
```

```{r eval=FALSE}
outputs$indicators %>%
  dplyr::filter(
    indicator == "prevalence",  # HIV prevalence
    age_group == "Y015_049"   # Age group 15-49
  ) %>%
  head()
```

```{r eval=FALSE}
add_output_labels(outputs) %>%
  dplyr::filter(
    indicator == "prevalence",  # HIV prevalence
    age_group == "Y015_049"   # Age group 15-49
  ) %>%
  head()
```

```{r eval=FALSE}
system.time(fit <- sample_tmb(fit))
```

```{r eval=FALSE}
system.time(outputs <- output_package(fit, naomi_mf))

outputs_calib <- calibrate_outputs(
  outputs, 
  naomi_mf,
  spectrum_plhiv_calibration_level = "national",
  spectrum_plhiv_calibration_strat = "sex_age_coarse",
  spectrum_artnum_calibration_level = "national", 
  spectrum_artnum_calibration_strat = "sex_age_coarse",
  spectrum_aware_calibration_level = "national", 
  spectrum_aware_calibration_strat = "sex_age_coarse",
  spectrum_infections_calibration_level = "national", 
  spectrum_infections_calibration_strat = "sex_age_coarse"
)

outputs$indicators %>%
  dplyr::filter(
    indicator == "prevalence",  # HIV prevalence
    age_group == "Y015_049"   # Age group 15-49
  ) %>%
  head()
```

# Using simplified TMB template

Use the same data from above, but fit a model using our simplified Naomi TMB template, rather than the one from the package.
This may require making some edits to the simplified TMB template, such as allowing non-integer values for survey-weighted cases and sample sizes.