### 2022-10-19 src/docs_xx-12-22-explainer/xx-12-22-explainer.Rmd
Template for lab meeting in December

---
title: Explaining approximate deterministic inference methods
subtitle: HIV Inference Lab Group Meeting
author: Adam Howes
institute: Imperial College London
date: December 2022
bibliography: citations.bib
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7 
    fig_height: 3
    includes:
        in_header: preamble.tex
---

# Bayes

* As a statistical modeller, our whole job (approximately) is to construct a generative model for data $y$ using parameters $\vartheta$
* This is the joint distribution $p(y, \vartheta) = p(y \, | \, \vartheta) p(\vartheta)$
* What we actually want is the posterior $p(\vartheta \, | \, y)$ which is **just**
$$
p(\vartheta \, | \, y) = \frac{p(y, \vartheta)}{p(y)} = \frac{p(y \, | \, \vartheta) p(\vartheta)}{p(y)}
$$
* The **core problem** of Bayesian inference is that we can't compute $p(y)$
$$
p(y) = \int p(y, \vartheta) \text{d} \vartheta
$$

# How might you do it?

* Usually if you wanted to integrate something, you could use numerical integration methods
* It doesn't work when $\text{dim}(\vartheta)$ gets even moderately sized

# Notation set-up

* Previously I had all of the parameters under the symbol $\vartheta$
* What if we split them up as being $\vartheta = (x, \theta)$

| Names for $x$ | Names for $\theta$ |
|---------------|--------------------|
| Latent variables, random effects, latent field | Hyperparameters, fixed effects

# Spatio-temporal statistics

* There is nothing inherently special about spatio-temporal statistics
* We have observations indexed by space $s \in \mathcal{S}$ and time $t \in \mathcal{T}$
* Usually we associate parameters to spatio-temporal locations as well as observations
* This ends up with us having something like $\{x_{s, t}\}$

# What's important about this?

1. There might be **a lot** of spatio-temporal locations, so $\text{dim}(x)$ might be pretty big!
 * If you have 100 districts and 10 years, that's already $100 \times 10 = 1000$ parameters
2. Perhaps we're willing to make assumptions about how things vary over space-time

<!-- As you get more data, the posterior approaches a Gaussian -->

# Latent Gaussian models

\begin{alignat*}{2}
  &\text{(Observations)}     &   y &\sim p(y \, | \, x, \theta), \\
  &\text{(Latent field)}     &   x &\sim \mathcal{N}(x  \, | \, \mu(\theta), Q(\theta)^{-1}), \\
  &\text{(Hyperparameters)}  &   \qquad \theta &\sim p(\theta),
\end{alignat*}
where $y = (y_1, \ldots, y_n)$, $x = (x_1, \ldots, x_n)$, $\theta = (\theta_1, \ldots, \theta_m)$

# References {.allowframebreaks}