### 2022-10-31 src/explore_posterior-comparison/posterior-comparison.Rmd
Set-up structure for posterior-comparison notebook

---
title: "Posterior comparison methods"
author:
- name: Adam Howes
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Background** Multiple approaches for conducting approximate Bayesian inference exist.

  **Task** We compare methods for assessing the accuracy of posterior distribution. These methods include marginal KS tests, simulation based calibration and maximum mean discrepancy.
---

# Data

# KS tests

# Simulation-based calibration

# Maximum mean discrepancy

# Original computing environment

```{r}
sessionInfo()
```