### 2022-11-10 src/docs_15-11-22-seminar/15-11-22-seminar.Rmd
Iterate slides forward

---
title: Integrated nested Laplace approximations for extended latent Gaussian models with application to the Naomi HIV model
subtitle: Waterloo SAS Student Seminar Series
author: Adam Howes
institute: Imperial College London
date: November 2022
bibliography: citations.bib
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7 
    fig_height: 3
    includes:
        in_header: preamble.tex
---

# Motivation

* Surveillance of the HIV epidemic in sub-Saharan Africa
* Want to estimate indicators used for monitoring and response, including:
  * Prevalence $\rho$: the proportion of people who are HIV positive
  * Incidence $\lambda$: the proportion of people newly infected 
  * Treatment coverage $\alpha$: the proportion of PLHIV on treatment
* We would like to provide them at a local, district-level

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]
This is a challenging task! Data is noisy, sparse and biased. $\implies$ compelling case for thoughtful Bayesian modelling.
\end{tcolorbox}
\end{center}

# A simple small-area model for prevalence

* Consider small-areas $i = 1, \ldots, n$ like the districts of a country
* Simple random sample household-survey of size $m^\text{HS}_i$ in each area
* The number of people testing positive for HIV is $y^\text{HS}_i$
* You could calculate direct estimates of prevalence by $y^\text{HS}_i / m^\text{HS}_i$ but because the survey is powered at a national-level, the sample sizes are small and these estimates would be noisy

# A simple small-area model for prevalence

* We can use a binomial logistic regression of the form:
\begin{align*}
y^\text{HS}_i &\sim \text{Bin}(m^\text{HS}_i, \rho^\text{HS}_i), \\
\text{logit}(\rho^\text{HS}_i) &\sim g(\vartheta^\text{HS}), \quad i = 1, \ldots, n,
\end{align*}
* We usually set up $g$ as a Gaussian spatial smoother
* This allows for pooling of information between districts

# Latent Gaussian models

* Three-stage Bayesian hierarchical model
\begin{alignat*}{2}
  &\text{(Observations)}     &   \y &\sim p(\y \, | \, \x), \\
  &\text{(Latent field)}     &   \x &\sim p(\x \, | \, \btheta), \\
  &\text{(Hyperparameters)}  &   \qquad \btheta &\sim p(\btheta),
\end{alignat*}
where $\y = (y_1, \ldots, y_n)$, $\x = (x_1, \ldots, x_n)$, $\btheta = (\theta_1, \ldots, \theta_m)$
* Interested in learning both $(\btheta, \x)$ from data $\y$
* If the middle layer is Gaussian, then it's a latent Gaussian model
\begin{equation*}
  \text{(Latent field)} \qquad  p(\x \, | \, \btheta) = \mathcal{N}(\x  \, | \, \bm{\mu}(\btheta), \bm{Q}(\btheta)^{-1}).
\end{equation*}
* Covers most of the models commonly used in spatiotemporal statistics
* Latent field is typically indexed by spatiotemporal location, such that $n > m$

# Limitations of household surveys

* A typical household survey takes around $UNKNOWN to run
* This means that they don't happen very often
* For this reason, the information might be quite out of date, and difficult to base policy on

# Adding ANC surveillance

* Pregnant women attending antenatal care clinics are routinely tested for HIV, to avoid mother-to-child transmission
* This data source is more real-time than household surveys, but it's also more biased, because attendees are unlikely to be as representative of the population
* But perhaps this bias is consistent, in which case we can still make use of the ANC data to supplement our model!

# Adding ANC surveillance

* Suppose of $m^\text{ANC}_i$ women attending ANC, $y^\text{ANC}_i$ are HIV positive, then we can use another binomial logistic regression:
\begin{align*}
y^\text{ANC}_i &\sim \text{Bin}(m^\text{ANC}_i, \rho^\text{ANC}_i), \\
\text{logit}(\rho^\text{ANC}_i) &= \text{logit}(\rho^\text{HS}_i) + b_i, \\
b_i &\sim \mathcal{N}(\beta_b, \sigma_b^2),
\end{align*}
* This is similar to using $\rho^\text{ANC}_i$ as a covariate in the model for household survey prevalence, but this way takes into account sampling variation

# Adding ART coverage

* We're also interested in what proportion $\alpha_i$ of people living with HIV (PLHIV) are receiving treatment
* Suppose we record $A_i$ attendees from a known population of $N_i$ in each district
* We can use another logistic regression model
\begin{align*}
A_i &\sim \text{Bin}(N_i, \rho^\text{HS}_i \alpha_i), \\
\text{logit}(\alpha_i) &\sim \mathcal{N}(\beta_\alpha, \sigma_\alpha^2).
\end{align*}

# Naomi evidence synthesis model

:::::::::::::: {.columns}

::: {.column width=.65}

* Combining these three modules is the basis of the Naomi evidence synthesis model
* Used by countries to produce HIV estimates in a yearly process supported by UNAIDS
* Can't run long MCMC in this setting, so we require fast, accurate, approximations 
* It's a complicated model, and requires something more flexible than `R-INLA`
* Currently using a package called Template Model Builder `TMB`

:::

::: {.column width=.35}

```{r, echo=FALSE, fig.cap="A supermodel", out.width = '65%'}
knitr::include_graphics("naomi_hex.png")
```

:::

::::::::::::::

#

```{r, echo=FALSE, fig.cap="Example of the user interface from https://naomi.unaids.org/", out.width = '70%'}
knitr::include_graphics("naomi_user.png")
```

# Template Model Builder

* `TMB` [@kristensen2015tmb] is an R package which implements the Laplace approximation for latent variable models
* To get started, write an objective function $f(\x, \btheta)$ in `TMB` C++ syntax
* As pseudo-Bayesians, we choose the log-posterior
$$
f(\x, \btheta) = - \log p(\y \, | \, \x, \btheta) p(\x \, | \, \btheta) p(\btheta)
$$

# Template Model Builder

For example, for the model
$$
\y \sim \mathcal{N}(\mu, 1)
$$
with $p(\mu) \propto 1$ then the `TMB` user template looks like...

#

```{cpp eval=FALSE}
#include <TMB.hpp>

template <class Type>
Type objective_function<Type>::operator()() {
  // Define data e.g.
  DATA_VECTOR(y);
  // Define parameters e.g.
  PARAMETER(mu);
  // Calculate negative log-likelihood e.g.
  nll = Type(0.0);
  nll -= dnorm(y, mu, 1, true).sum()
  return(nll);
}
```

# Template Model Builder

* We can use `TMB` to obtain the Laplace approximation
$$
\tilde p_\text{LA}(\btheta \, | \, \y) \propto \frac{p(\y, \x, \btheta)}{\tilde p_\text{G}(\x \, | \, \btheta, \y)} \Big\rvert_{\x = \bm{\mu}^\star(\btheta)} \label{eq:hypermarginal}
$$
* Integrate out a Gaussian approximation $\tilde p_\text{G}(\x \, | \, \btheta, \y)$ to the latent field
* `TMB` uses automatic differentiation [@griewank2008evaluating] via `CppAD`

# Integrated Nested Laplace Approximation

* Integrated nested Laplace approximation (INLA) [@rue2009approximate; @blangiardo2015spatial] is an approach to approximate inference which builds on the Laplace approximation
* Goal is to approximate \textcolor{hilit}{posterior marginals} $\{\tilde p(x_i \, | \, \y)\}_{i = 1}^n$ and $\{\tilde p(\theta_j \, | \, \y)\}_{j = 1}^m$
\begin{align}
  p(x_i \, | \, \y) &= \int p(x_i, \btheta \, | \, \y) \text{d} \btheta = \int p(x_i \, | \, \btheta, \y) p(\btheta \, | \, \y) \text{d}\btheta, \quad i = 1, \dots, n, \label{eq:inla1} \\
  p(\theta_j \, | \, \y) &= \int p(\btheta \, | \, \y) \text{d}\btheta_{-j} \quad j = 1, \ldots, m. \label{eq:inla2}
\end{align}
* To do so, we require the approximations $\tilde p(\btheta \, | \, \y)$ and $\tilde p(x_i \, | \, \btheta, \y)$
* There are four steps as to how the method works (bare with me!)

# Step 1)

1) First Laplace approximate hyperparameter posterior
\begin{equation}
\tilde p_\text{LA}(\btheta \, | \, \y) \propto \frac{p(\y, \x, \btheta)}{\tilde p_\text{G}(\x \, | \, \btheta, \y)} \Big\rvert_{\x = \bm{\mu}^\star(\btheta)} \label{eq:hypermarginal}
\end{equation}
which can be marginalised to get $\tilde p(\theta_j \, | \, \y)$
* Notice that this is the same object we had been working with in `TMB`
* We use this approximation \textcolor{hilit}{nested} within integrals -- hence the name INLA

# Step 2)

2) In both Equations \eqref{eq:inla1} and \eqref{eq:inla2} we want to integrate w.r.t. $\btheta$, so choose integration nodes and weights $\{ \btheta(\z), \omega(\z) \}_{\z \in \mathcal{Z}}$
  * For low $m$ `R-INLA` uses a grid-strategy
  * For larger $m$ this becomes too expensive and `R-INLA` uses a CCD design
  * We plan to use adaptive Gaussian Hermite quadrature (AGHQ), which has recently been shown to have theoretical guarantees [@bilodeau2021stochastic]

#

```{r message=FALSE, echo=FALSE, out.height="200px", fig.cap=paste("An illustration of the \\texttt{R-INLA} grid method for selecting integration nodes using a toy bivariate Gaussian distribution for $\\btheta$. Start at the mode and work outwards along the eigenvectors until the density drops sufficiently low. \\label{fig:grid}")}
knitr::include_graphics("depends/inla-grid.pdf")
```

# Adaptive Gaussian Hermite Quadrature

* Gauss-Hermite quadrature is a way of picking nodes and weights, and is based on the theory of polynomial interpolation
* The adaptive part means that it uses the location (mode) and curvature (Hessian) of the target (posterior) to automatically choose the node locations
  * Does not require manual tuning!
* Works particularly well when the integrand is pretty Gaussian
* Use $k$ quadrature nodes per dimension, for example if $k = 3$ then $3^m$ total nodes
* Implemented in the `aghq` R package. See vignette @stringer2021implementing

#

```{r message=FALSE, echo=FALSE, out.height="200px", fig.cap=paste("One dimensional example of AGHQ. \\label{fig:aghq-1d}")}
knitr::include_graphics("depends/aghq-1d.pdf")
```

#

```{r message=FALSE, echo=FALSE, out.height="200px", fig.cap=paste("Two dimensional example of AGHQ. \\label{fig:aghq-2d}")}
knitr::include_graphics("depends/aghq-2d.pdf")
```

# Step 3)

3) Choose approximation for $\tilde p(x_i \, | \, \btheta, \y)$
  * Simplest version [@rue2007approximate] is to marginalise $\tilde p_\text{G}(\x \, | \, \btheta, \y)$
\begin{equation}
\tilde p_\text{G}(x_i \, | \, \btheta, \y) = \mathcal{N}(x_i \, | \, \mu^\star_i(\btheta), 1 / q^\star_i(\btheta))
\end{equation}
  * In `R-INLA`, the above is referred to as `method = "gaussian"` 
  * There are two better, more complex approximations, confusingly called `"simplified.laplace"` and `"laplace"`
  * Uses sparsity properties of $\bm{Q}(\btheta)$, i.e. if $\x$ is a Gaussian Markov random field (GMRF)

# Step 4)

4) Finally, use quadrature to combine
  * our approximation $\tilde p_\text{LA}(\btheta \, | \, \y)$ from step 1),
  * some choice of integration nodes and weights $\{ \btheta(\z), \omega(\z) \}$ from step 2),
  * some choice of approximation $\tilde p(x_i \, | \, \btheta, \y)$ from step 3)
to give
\begin{equation}
  \tilde p(x_i \, | \, \y) = 
  \sum_{\z \in \mathcal{Z}} \tilde p(x_i \, | \, \btheta(\z), \y) \times \tilde p_\text{LA}(\btheta(\z) \, | \, \y) \times \omega(\z)
\end{equation}

# Experiments

* We wrote a simplified version of the Naomi model up in `TMB`
* This allowed us to test three inference methods **all using precisely the same model and C++ code**

1. A direct Gaussian approximation via `TMB`
2. Adaptive Gaussian Hermite quadrature via `aghq`
3. No-U-Turn Sampling (NUTS -- a type of Hamiltonian Monte Carlo) via `tmbstan`

* Note: using different software it is usually very difficult to ensure the model is precisely the same, so we're very fortunate here

# Comparison approach

* You could look at the summaries like the mean and standard deviation of each of the posterior marginals
  * Any approximation method should be pretty good at getting the mean right
  * Gaussian approximations should be good at getting the second moment right
* It's probably better to compare the whole posterior distributions
* One way to do this is via Kolmogorov-Smirnov statistics, which give the maximum difference between two empirical CDFs

# References {.allowframebreaks}