### 2022-11-21 src/docs_bayescomp-poster/bayescomp-poster.Rmd
Small poster update

---
main_topsize: 0.18
main_bottomsize: 0.03
poster_height: "48in"
poster_width: "36in"
font_family: "Open Sans"
# Essentials
title: Integrated nested Laplace approximations for extended latent Gaussian models, with application to the Naomi HIV model
author:
  - name: '**Adam Howes**'
    affil: 1, 2
    main: true
    twitter: adamhowes
    email: ath19@ic.ac.uk
  - name: "Alex Stringer"
    affil: 3
  - name: "Seth R. Flaxman"
    affil: 4
  - name: "Jeffrey W. Eaton"
    affil: 2
affiliation:
  - num: 1
    address: Department of Mathematics, Imperial College London
  - num: 2
    address: MRC Centre for Global Infectious Disease Analysis, School of Public Health, Imperial College London
  - num: 3
    address: Department of Statistics and Actuarial Science, University of Waterloo
  - num: 4
    address: Department of Computer Science, University of Oxford
main_findings: "Simplified INLA, automatic differentiation and adaptive Gauss-Hermite quadrature for fast and accurate approximate inference"
logoleft_name: "IMP_ML_W_N.png"
# logoright_name: "UNAIDS-reverse-EN.gif"
# Style & formatting
column_numbers: 3
title_textsize: "60px"
author_textsize: "1.17em"
authorextra_textsize: "35px"
affiliation_textsize: "25px"
affiliation_textcol: '#00000080'
caption_fontsize: "20px"
# Main
main_fontfamily: "Open Sans"
# Poster body options
primary_colour: '#38761C'
secondary_colour: '#3b77D8'
accent_colour: "#5d19c4"
body_bgcol: "#ffffff"
body_textsize: "35px"
body_textcol: "#000000"
reference_textsize: "25px"
sectitle_textsize: "40px"
# Standard options
bibliography: citations.bib
output: 
  posterdown::posterdown_betterport:
    self_contained: false
    pandoc_args: --mathjax
    highlight: espresso
    number_sections: false
---

<!-- Using LaTeX to create spacing -->
$$\\[0.15in]$$

## The Naomi model

$$\\[0.15in]$$

## Inference procedure

### Laplace approximation

### Adaptive Gauss-Hermite Quadrature

Approximate integrals by 
$$
\int_\Theta p(\theta) \text{d} \theta \approx \lvert L \rvert \sum_{z \in \mathcal{Q}(m, k)} p(\hat \theta + Lz) \omega(z)
$$
with **Gauss-Hermite quadrature** rule $z \in \mathcal{Q}(m, k)$ **adapted** based upon the mode $\hat \theta = \text{argmax}_{\theta \in \Theta} \in p(\theta)$ and lower Cholesky $LL^\top = - \partial_\theta^2 \log p(\theta) \rvert_{\theta = \hat \theta}$ of the target.

### Our algorithm

Given C++ user template for $- \log p(y, x, \theta)$:

## Comparison

## Conclusions

<font size="6"> **Funding** AH was supported by the EPSRC and Bill & Melinda Gates Foundation. This research was supported by the MRC Centre for Global Infectious Disease Analysis. </font>

## References

<div id="refs" class="references"></div>