### 2022-12-05 src/example_naomi/maths.Rmd
More maths writing

---
title: "Mathematical description of Naomi"
author:
- name: Adam Howes
output:
  html_document:
    toc: yes
    toc_float: true
    toc_collapsed: true
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
    **Background** The Naomi HIV model is specified in the appendix of Eaton et al. (2021).

    **Task** We translate the model to R Markdown in order to (1) aid understanding, (2) clarify any details which have been compressed for publication, (3) avoid having to read mathematics in Word.
---

# Notation

* District $x$
  * Located in Spectrum region file $R_x$
* Sex $s \in \{\text{F}, \text{M}\}$
* Age $a \in \{\text{0-5}, \text{5-10}, \ldots, \text{75-80}, 80+\}$
* Time $t \in \{T_1, T_2, T_3\}$
  * $T_1$: Most recent national household survey with HIV testing
  * $T_2$: Current time period at which to generate estimates
  * $T_3$: Short-term project period (typically 9-12 months)
  * Note $T_{1:3}$ specified to nearest calendar quarter
* Population size $N_{x, s, a, t}$
* HIV prevalence $\rho_{x, s, a, t}$
* ART coverage $\alpha_{x, s, a, t}$
* Annual HIV incidence rate $\lambda_{x, s, a, t}$

# HIV prevalence and ART coverage at $T_1$

## Background

* $u \sim \text{ICAR}(\sigma)$ refers to the intrinsic conditional auto-regressive model (ICAR) with marginal standard deviation $\sigma > 0$.
* $u \sim \text{BYM}2(\sigma, \phi)$ refers to a reparameterised Besag-York-Mollie model (BYM2) with marginal standard deviation $\sigma > 0$ comprised of a spatially structured ICAR component with proportion $\phi \in (0, 1)$ and spatially unstructured IID component with proportion $1 - \phi$.
* For both the ICAR and BYM2 models, we follow recommendations on scaling, disconnected adjacency graph components, and islands.

## HIV prevalence

HIV prevalence modelled by
$$
\text{logit}(\rho_{x, s, a, T_1}) = \beta^\rho_0 + \beta_{S}^{\rho, s = \text{M}} + u^\rho_a + u_a^{\rho, s = \text{M}} + u^\rho_x + u_x^{\rho, s = \text{M}} + u_x^{\rho, a < 15} + \eta^\rho_{R_x, s, a} 
$$
where:

* $\beta^\rho_0$ is the intercept
* $\beta_{s}^{\rho, s = \text{M}}$ is the difference in logit prevalence for men compared to women
* $u^\rho_a \sim \text{AR}1(\sigma_A^\rho, \phi_A^\rho)$ are age random effects for women
* $u_a^{\rho, s = \text{M}}$ are age random effects for the difference in logit prevalence for men compared to women age $a$
* $u^\rho_x \sim \text{BYM}2(\sigma_X^\rho, \phi_X^\rho)$ are spatial random effects for women
* $u_x^{\rho, s = \text{M}} \sim \text{BYM}2(\sigma_{XS}^\rho, \phi_{XS}^\rho)$ are spatial random effects for the difference in logit prevalence for men compared to women in district $x$
* $u_x^{\rho, a < 15} \sim \text{ICAR}(0, \sigma_{XA}^\rho)$ are spatial random effects for the ratio of paediatric prevalence to adult women prevalence
* $\eta^\rho_{R_x, s, a}$ are a fixed offset specifying assumed odds ratios for prevalence outside the age ranges for which data are available

For prior distributions, we use:

* $\mathcal{N}(0, 5)$ for all fixed effects ($\beta^\rho_0, \beta_{s}^{\rho, s = \text{M}}$)
* $\mathcal{N}^{+}(0, 2.5)$ for all standard deviation terms
* $\mathcal{U}(-1, 1)$ for all AR1 correlation parameters
* $\text{Beta}(0.5, 0.5)$ for all BYM2 proportion parameters

## ART coverage

ART coverage modelled by
$$
\text{logit}(\alpha_{x, s, a, T_1}) = \beta^\alpha_0 + \beta_{S}^{\alpha, s = \text{M}} + u^\alpha_a + u_a^{\alpha, s = \text{M}} + u^\alpha_x + u_x^{\alpha, s = \text{M}} + u_x^{\alpha, a < 15} + \eta^\alpha_{R_x, s, a} 
$$
where terms are analogous to the HIV prevalence model.

## HIV incidence rate

HIV incidence rate is modelled by
$$
\log(\lambda_{x, s, a, t}) = \beta_0^\lambda + \beta_S^{\lambda, s = \text{M}} + \log(\rho_{x, t}^{\text{15-49}}) + \log(1 - \omega \cdot \alpha_{x, t}^{\text{15-49}}) + u_x^\lambda + \eta_{R_x, s, a, t}^\lambda 
$$
where:

* $\beta^\lambda_0$ is the intercept, which is proportional to the average HIV transmission rate for untreated HIV positive adults
* $\beta_S^{\lambda, s = \text{M}}$ is the log incidence rate ratio for men compared to women
* $\rho_{x, t}^{\text{15-49}}$ is the HIV prevalence among adults 15-49 calculated by
$$
\rho_{x, t}^{\text{15-49}} = \frac{\sum_{s \in \{\text{F}, \text{M}\}} \sum_{a = 15}^{49} N_{x, s, a, t} \cdot \rho_{x, s, a, t}}{\sum_{s \in \{\text{F}, \text{M}\}} \sum_{a = 15}^{49} N_{x, s, a, t}}
$$
* $\alpha_{x, t}^{\text{15-49}}$ is the ART coverage among adults 15-49 calculated by
$$
\alpha_{x, t}^{\text{15-49}} = \frac{\sum_{s \in \{\text{F}, \text{M}\}} \sum_{a = 15}^{49} N_{x, s, a, t} \cdot \rho_{x, s, a, t} \cdot \alpha_{x, s, a, t}}{\sum_{s \in \{\text{F}, \text{M}\}} \sum_{a = 15}^{49} N_{x, s, a, t} \cdot \rho_{x, s, a, t}}
$$
* $\omega$ is the average reduction in HIV transmission rate per 1% increase in population ART coverage and is fixed at $\omega = 0.7$ from the EPP model
* $u_x^\lambda \sim \mathcal{N}(0, \sigma^\lambda)$ with $\sigma^\lambda \sim \mathcal{N}^+(0, 1)$ are IID spatial random effects
* $\eta^\lambda_{R_x, s, a, t}$ specify log incidence rate ratios by sex and age group calculated from Spectrum model output

## Short-term projection from $T_1 \to T_2$ and $T_2 \to T_3$

### HIV prevalence

* HIV population projected from $T_1 \to T_2$ based on survival of PLHIV from Spectrum output, ageing from one five-year age group to the next, and the addition of new HIV infections calculated from district-level HIV incidence rate and survival and ageing after infection
* Let $H_{x, s, a, {T_1}} = N_{x, s, a, {T_1}} \cdot \rho_{x, s, a, {T_1}}$ be the number of PLHIV at $T_1$
* For age groups $a \geq 5$ the number of PLHIV at $T_2$ is modelled by
$$
\begin{align*}
H_{x, s, a, {T_2}} &= H_{x, s, a, {T_1}} \cdot S_{{R_x}, s, a \to a, {T_1}} + H_{x, s, a - 5, {T_1}} \cdot S_{{R_x}, s, a - 5 \to a, {T_1}} \\
&+ (1 - \exp(- \delta_{T_1} \cdot \lambda_{x, s, a, {T_1}})) \cdot (N_{x, s, a, {T_1}} - H_{x, s, a, {T_1}}) \cdot L_{R_x, s, a \to a, {T_1}} \\
&+ (1 - \exp(- \delta_{T_1} \cdot \lambda_{x, s, a - 5, {T_1}})) \cdot (N_{x, s, a - 5, {T_1}} - H_{x, s, a - 5, {T_1}}) \cdot L_{R_x, s, a - 5 \to a, {T_1}}
\end{align*}
$$
* $S_{{R_x}, s, a \to a, {T_1}}$ is the ratio for the number of PLHIV surviving from $T_1$ to $T_2$ and remaining in age group $a$ (i.e. not aging from $a$ to $a + 5$)
* $S_{{R_x}, s, a - 5 \to a, {T_1}}$ the survival ratio from $T_1$ to $T_2$ and aging from $a - 5$ to $a$
* $(1 - \exp(- \delta_{T_1} \cdot \lambda_{x, s, a, {T_1}}))$ is the probability of acquiring HIV between $T_1$ and $T_2$ of duration $\Delta_{T_1} = T_2 - T_1$ for the susceptible population $N_{x, s, a, {T_1}} - H_{x, s, a, {T_1}}$
* $L_{R_x, s, a \to a, {T_1}}$ is the ratio surviving and remaining in age group $a$ for those infected between $T_1$ and $T_2$
* $L_{R_x, s, a - 5 \to a, {T_1}}$ is the ratio surviving and aging from $a - 5$ to $a$ for those infected between $T_1$ and $T_2$
* The ratios $S$ and $L$ are calculated from Specturm results by
  1. Disaggregating single-year or single-age Specturm model results for PLHIV at $T_1$ and $T_2$ and new infections between $T_1$ and $T_2$ to quarterly birth cohorts
  2. Subtracting the number of new infections within the cohort from the number of PLHIV at $T_2$ to calculate the number of surviving PLHIV in each cohort
  3. Aggregating the survivors by quarter-age cohorts to those who survived and aged from $a$ to $a'$ between $T_1$ and $T_2$, and calculating the ratio by dividing by the initial PLHIV $H_{x, s, a, {T_1}}$
* Variation in ART coverage or effectiveness is assumed not to affect survival

### ART coverage

* ART coverage is modelled by
$$
\text{logit}(\alpha_{x, s, a, {T_2}}) = \text{logit}(\alpha_{x, s, a, {T_1}}) + \beta_{T_2}^\alpha+\beta_{T_2, s=M}^\alpha+u_{x, T 2}^\alpha+u_{x, a<15, T_2}^\alpha+\eta_{R_x, s, a, T 2}^\alpha
$$

## Clarification notes to discuss with Jeff

* Some confusion on how to offset terms work -- they seem cool though. Might prefer they not be called $\eta$ which is in my head linear predictor
  * $\eta^\lambda_{R_x, s, a, t}$ aren't these $R_x, s, a, t$ specific rather than 
* Aggregation over age 15-49 uses sum from $a = 15$ to $a = 45$? Perhaps this is because $a = 45$ actually corresponds to the age band 45-49
* Sometimes in subscript we have $x, s, a, t$ and sometimes e.g. $x, s, at$ -- why?
* $\sigma^\lambda \sim \mathcal{N}^+(0, 1)$ rather than e.g. $\sigma^\lambda \sim \mathcal{N}^+(0, 2.5)$
* Inconsistent use of $T2$ and $T_2$