### 2022-12-06 src/example_naomi/maths.Rmd
Wow, more maths!

---
title: "Mathematical description of Naomi"
author:
- name: Adam Howes
output:
  html_document:
    toc: yes
    toc_float: true
    toc_collapsed: true
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
    **Background** The Naomi HIV model is specified in the appendix of Eaton et al. (2021).

    **Task** We translate the model to R Markdown in order to (1) aid understanding, (2) clarify any details which have been compressed for publication, (3) avoid having to read mathematics in Word.
---

# Process specification

## Notation

* District $x$
  * Located in Spectrum region file $R_x$
* Sex $s \in \{\text{F}, \text{M}\}$
* Age $a \in \{\text{0-5}, \text{5-10}, \ldots, \text{75-80}, 80+\}$
* Time $t \in \{T_1, T_2, T_3\}$
  * $T_1$: Most recent national household survey with HIV testing
  * $T_2$: Current time period at which to generate estimates
  * $T_3$: Short-term project period (typically 9-12 months)
  * Note $T_{1:3}$ specified to nearest calendar quarter
* Population size $N_{x, s, a, t}$
* HIV prevalence $\rho_{x, s, a, t}$
* ART coverage $\alpha_{x, s, a, t}$
* Annual HIV incidence rate $\lambda_{x, s, a, t}$
* $u \sim \text{ICAR}(\sigma)$ refers to the intrinsic conditional auto-regressive model (ICAR) with marginal standard deviation $\sigma > 0$.
* $u \sim \text{BYM}2(\sigma, \phi)$ refers to a reparameterised Besag-York-Mollie model (BYM2) with marginal standard deviation $\sigma > 0$ comprised of a spatially structured ICAR component with proportion $\phi \in (0, 1)$ and spatially unstructured IID component with proportion $1 - \phi$.
* For both the ICAR and BYM2 models, we follow recommendations on scaling, disconnected adjacency graph components, and islands.

## HIV prevalence and ART coverage at $T_1$

### HIV prevalence

HIV prevalence modelled by
$$
\text{logit}(\rho_{x, s, a, T_1}) = \beta^\rho_0 + \beta_{S}^{\rho, s = \text{M}} + u^\rho_a + u_a^{\rho, s = \text{M}} + u^\rho_x + u_x^{\rho, s = \text{M}} + u_x^{\rho, a < 15} + \eta^\rho_{R_x, s, a} 
$$
where:

* $\beta^\rho_0$ is the intercept
* $\beta_{s}^{\rho, s = \text{M}}$ is the difference in logit prevalence for men compared to women
* $u^\rho_a \sim \text{AR}1(\sigma_A^\rho, \phi_A^\rho)$ are age random effects for women
* $u_a^{\rho, s = \text{M}}$ are age random effects for the difference in logit prevalence for men compared to women age $a$
* $u^\rho_x \sim \text{BYM}2(\sigma_X^\rho, \phi_X^\rho)$ are spatial random effects for women
* $u_x^{\rho, s = \text{M}} \sim \text{BYM}2(\sigma_{XS}^\rho, \phi_{XS}^\rho)$ are spatial random effects for the difference in logit prevalence for men compared to women in district $x$
* $u_x^{\rho, a < 15} \sim \text{ICAR}(0, \sigma_{XA}^\rho)$ are spatial random effects for the ratio of paediatric prevalence to adult women prevalence
* $\eta^\rho_{R_x, s, a}$ are a fixed offset specifying assumed odds ratios for prevalence outside the age ranges for which data are available

For prior distributions, we use:

* $\mathcal{N}(0, 5)$ for all fixed effects ($\beta^\rho_0, \beta_{s}^{\rho, s = \text{M}}$)
* $\mathcal{N}^{+}(0, 2.5)$ for all standard deviation terms
* $\mathcal{U}(-1, 1)$ for all AR1 correlation parameters
* $\text{Beta}(0.5, 0.5)$ for all BYM2 proportion parameters

### ART coverage

ART coverage modelled by
$$
\text{logit}(\alpha_{x, s, a, T_1}) = \beta^\alpha_0 + \beta_{S}^{\alpha, s = \text{M}} + u^\alpha_a + u_a^{\alpha, s = \text{M}} + u^\alpha_x + u_x^{\alpha, s = \text{M}} + u_x^{\alpha, a < 15} + \eta^\alpha_{R_x, s, a} 
$$
where terms are analogous to the HIV prevalence model.

### HIV incidence rate

HIV incidence rate is modelled by
$$
\log(\lambda_{x, s, a, t}) = \beta_0^\lambda + \beta_S^{\lambda, s = \text{M}} + \log(\rho_{x, t}^{\text{15-49}}) + \log(1 - \omega \cdot \alpha_{x, t}^{\text{15-49}}) + u_x^\lambda + \eta_{R_x, s, a, t}^\lambda 
$$
where:

* $\beta^\lambda_0$ is the intercept, which is proportional to the average HIV transmission rate for untreated HIV positive adults
* $\beta_S^{\lambda, s = \text{M}}$ is the log incidence rate ratio for men compared to women
* $\rho_{x, t}^{\text{15-49}}$ is the HIV prevalence among adults 15-49 calculated by
$$
\rho_{x, t}^{\text{15-49}} = \frac{\sum_{s \in \{\text{F}, \text{M}\}} \sum_{a = 15}^{49} N_{x, s, a, t} \cdot \rho_{x, s, a, t}}{\sum_{s \in \{\text{F}, \text{M}\}} \sum_{a = 15}^{49} N_{x, s, a, t}}
$$
* $\alpha_{x, t}^{\text{15-49}}$ is the ART coverage among adults 15-49 calculated by
$$
\alpha_{x, t}^{\text{15-49}} = \frac{\sum_{s \in \{\text{F}, \text{M}\}} \sum_{a = 15}^{49} N_{x, s, a, t} \cdot \rho_{x, s, a, t} \cdot \alpha_{x, s, a, t}}{\sum_{s \in \{\text{F}, \text{M}\}} \sum_{a = 15}^{49} N_{x, s, a, t} \cdot \rho_{x, s, a, t}}
$$
* $\omega$ is the average reduction in HIV transmission rate per 1% increase in population ART coverage and is fixed at $\omega = 0.7$ from the EPP model
* $u_x^\lambda \sim \mathcal{N}(0, \sigma^\lambda)$ with $\sigma^\lambda \sim \mathcal{N}^+(0, 1)$ are IID spatial random effects
* $\eta^\lambda_{R_x, s, a, t}$ specify log incidence rate ratios by sex and age group calculated from Spectrum model output

## Short-term projection from $T_1 \to T_2$ and $T_2 \to T_3$

### HIV prevalence

* HIV population projected from $T_1 \to T_2$ based on survival of PLHIV from Spectrum output, ageing from one five-year age group to the next, and the addition of new HIV infections calculated from district-level HIV incidence rate and survival and ageing after infection
* Let $H_{x, s, a, {T_1}} = N_{x, s, a, {T_1}} \cdot \rho_{x, s, a, {T_1}}$ be the number of PLHIV at $T_1$
* For age groups $a \geq 5$ the number of PLHIV at $T_2$ is modelled by
$$
\begin{align*}
H_{x, s, a, {T_2}} &= H_{x, s, a, {T_1}} \cdot S_{{R_x}, s, a \to a, {T_1}} + H_{x, s, a - 5, {T_1}} \cdot S_{{R_x}, s, a - 5 \to a, {T_1}} \\
&+ (1 - \exp(- \delta_{T_1} \cdot \lambda_{x, s, a, {T_1}})) \cdot (N_{x, s, a, {T_1}} - H_{x, s, a, {T_1}}) \cdot L_{R_x, s, a \to a, {T_1}} \\
&+ (1 - \exp(- \delta_{T_1} \cdot \lambda_{x, s, a - 5, {T_1}})) \cdot (N_{x, s, a - 5, {T_1}} - H_{x, s, a - 5, {T_1}}) \cdot L_{R_x, s, a - 5 \to a, {T_1}}
\end{align*}
$$
* $S_{{R_x}, s, a \to a, {T_1}}$ is the ratio for the number of PLHIV surviving from $T_1$ to $T_2$ and remaining in age group $a$ (i.e. not aging from $a$ to $a + 5$)
* $S_{{R_x}, s, a - 5 \to a, {T_1}}$ the survival ratio from $T_1$ to $T_2$ and aging from $a - 5$ to $a$
* $(1 - \exp(- \delta_{T_1} \cdot \lambda_{x, s, a, {T_1}}))$ is the probability of acquiring HIV between $T_1$ and $T_2$ of duration $\Delta_{T_1} = T_2 - T_1$ for the susceptible population $N_{x, s, a, {T_1}} - H_{x, s, a, {T_1}}$
* $L_{R_x, s, a \to a, {T_1}}$ is the ratio surviving and remaining in age group $a$ for those infected between $T_1$ and $T_2$
* $L_{R_x, s, a - 5 \to a, {T_1}}$ is the ratio surviving and aging from $a - 5$ to $a$ for those infected between $T_1$ and $T_2$
* The ratios $S$ and $L$ are calculated from Specturm results by
  1. Disaggregating single-year or single-age Specturm model results for PLHIV at $T_1$ and $T_2$ and new infections between $T_1$ and $T_2$ to quarterly birth cohorts
  2. Subtracting the number of new infections within the cohort from the number of PLHIV at $T_2$ to calculate the number of surviving PLHIV in each cohort
  3. Aggregating the survivors by quarter-age cohorts to those who survived and aged from $a$ to $a'$ between $T_1$ and $T_2$, and calculating the ratio by dividing by the initial PLHIV $H_{x, s, a, {T_1}}$
* Variation in ART coverage or effectiveness is assumed not to affect survival

### ART coverage

Change in ART coverage is modelled by
$$
\text{logit}(\alpha_{x, s, a, {T_2}}) = \text{logit}(\alpha_{x, s, a, {T_1}}) + \beta_{T_2}^\alpha + \beta_{T_2, s = \text{M}}^\alpha + u_{x, {T_2}}^\alpha + u_{x, a < 15, {T_2}}^\alpha + \eta_{R_x, s, a, {T_2}}^\alpha
$$
where

* $\eta_{R_x, s, a, {T_2}}^\alpha$ are offsets for the logit change in ART coverage between $T_1$ and $T_2$ from Spectrum region $R_x$
* $\beta_{T_2}^\alpha$ is the average change in logit ART coverage between $T_1$ and $T_2$ for adult women
* $\beta_{T_2, s = \text{M}}^\alpha$ is the difference in average change for men compared to women
* $u_{x, {T_2}}^\alpha \sim \mathcal{N}(0, \sigma_{XT}^\alpha)$ are district random effects
* $u_{x, a < 15, {T_2}}^\alpha$ is the difference in logit ART coverage for children compared to adults

**Flag for not understanding** ART coverage at $T_3$ is modelled by 
$$
\text{logit}(\alpha_{x, s, a, {T_2}}) = \text{logit}(\alpha_{x, s, a, {T_1}}) + \eta_{R_x, s, a, {T_3}}^\alpha
$$

## ANC testing cascade

* For women 15-49 the predicted number of ANC clients $\Psi_{x, a, t}$ is a log-linear model
$$
\log (\Psi_{x, a, t}) = \log (N_{x, \text{F}, a, t}) + \psi_{R_x, a, t} + \beta^\psi + u_x^\psi
$$
  * $N_{x, \text{F}, a, t}$ are the female population sizes
  * $\psi_{R_x, a, t}$ are ASFR in Spectrum region $R_x$ at time $t$
  * $\beta^\psi$ are the log rate ratio for the number of ANC clients relative to the predicted fertility
  * $u_x^\psi \sim \mathcal{N}(0, \sigma^\psi)$ are district random effects
* HIV prevalence $\rho_{x, a, t}^\text{ANC}$ and ART coverage $\alpha_{x, a, t}^\text{ANC}$ among pregnant women modelled with logit-linear models
\begin{align}
\text{logit}(\rho_{x, a, t}^{\text{ANC}}) &= \text{logit}(\rho_{x, F, a, t}) + \beta^{\rho^{\text{ANC}}} + \beta_{t \in(T_2, T_3\}}^{\rho^{\text{ANC}}} + u_x^{\rho^{\text{ANC}}} + u_{x, t \in\{T_2, T_3\}}^{\rho^{\text{ANC}}} + \eta_{R_x, a, t}^{\rho^{\text{ANC}}} \\
\text{logit}(\alpha_{x, a, t}^{\text{ANC}}) &= \text{logit}(\alpha_{x, F, a, t}) + \beta^{\alpha^{\text{ANC}}} + \beta_{t \in\{T_2, T_3\}}^{\alpha^{\text{ANC}}} + u_x^{\alpha^{\text{ANC}}}+u_{x, t \in\{T_2, T_3\}}^{\alpha^{\text{ANC}}} + \eta_{R_x, a, t}^{\alpha^{\text{ANC}}}
\end{align}
  * $\eta_{R_x, a, t}^{\theta^{\text{ANC}}}$ for $\theta \in \{\rho, \alpha\}$ are offsets for the log fertility rate ratios for HIV positive women compared to HIV negative women and for women on ART to HIV positive women not on ART, calculated from Spectrum model outputs for region $R_x$
  * $\beta^{\theta^{\text{ANC}}}$ for $\theta \in \{\rho, \alpha\}$ are the average differences between population and ANC outcomes at $T_1$ after removing the offset
  * $\beta_{t \in(T_2, T_3\}}^{\theta^{\text{ANC}}}$ for $\theta \in \{\rho, \alpha\}$ are the change in average difference from $T_1$ to $T_2$
  * $u_x^{\theta^{\text{ANC}}} \sim \mathcal{N}(0, \sigma_X^{\theta^{\text{ANC}}})$ for $\theta \in \{\rho, \alpha\}$ are district random effects
  * $u_{x, t \in\{T_2, T_3\}}^{\theta^{\text{ANC}}} \sim \mathcal{N}(0, \sigma_{XT}^{\theta^{\text{ANC}}})$ for $\theta \in \{\rho, \alpha\}$ capture the change in district random effects between times $T_1$ and $T_2$

## ANC attendance

* Number of PLHIV on ART is $A_{x, s, a, t} = N_{x, s, a, t} \cdot \rho_{x, s, a, t} \cdot \alpha_{x, s, a, t}$
* Let $\gamma_{x, x', t} \in [0, 1]$ be the probability that a person of ART residing in district $x$ recieves ART in district $x'$ at time $t$
* We assume that $\gamma_{x, x', t} = 0$ for $x \notin \{x, \text{ne}(x)\}$ i.e. that individuals seek treatment only in their residing district and its neighbours $\text{ne}(x) = \{x': x' \sim x\}$ where $\sim$ is an adjacency relation
  * $\sum_{x' \in \{x, \text{ne}(x)\}} \gamma_{x, x', t} = 1$
* Model $\gamma_{x, x', t}$ with a multinomial logit model where $\tilde \gamma_{x, t}$ is the log odds ratio of seeking ART in the district $x'$ neighbouring $x$ **flag for confusion here**
\begin{align}
\tilde \gamma_{x, t} &= \tilde \gamma_0 + u_x^{\tilde \gamma} + u_x^{\tilde \gamma, t = T_2} \\ 
\tilde \gamma_0 &= -4 \\
u_x^{\tilde \gamma} &\sim \mathcal{N}(0, \sigma_X^{\tilde \gamma}) \\
u_x^{\tilde \gamma, t \in \{T_2, T_3\}} &\sim \mathcal{N}(0, \sigma_{XT}^{\tilde \gamma}) 
\end{align}
  * $\tilde \gamma_{x, t}$ is an intercept specifying the prior mean for the log odds of seeking ART in each neighbouring district compared to the home district
  * $u_x^{\tilde \gamma}$ are district random effects
  * $u_x^{\tilde \gamma, t \in \{T_2, T_3\}}$ is the change in district level log-odds ratios between $T_1$ and $T_2$ -- **flag for confusion: shouldn't this just be $t = T_2$ here**
  * Note: only one value estimated for each district, implying no preference between neighbouring districts, and no sex-age specific modelling
* Calculate multinomial probabilities via softmax (n.b. this is the sigmoid function, the inverse of the softmax)
$$
\gamma_{x, x', t} = \frac{1}{1 + \sum_{x^\star \sim x} \exp(\tilde \gamma_{x^\star, t})}
$$
* The number of ART clients who reside in district $x$ and obtain ART in district $x'$ are $A_{x, x', s, a, t} = A_{x, s, a, t} \cdot \gamma_{x, x', t} = N_{x, s, a, t} \cdot \rho_{x, s, a, t} \cdot \alpha_{x, s, a, t} \cdot \gamma_{x, x', t}$
* Total attending ART facilities in district $x'$ is $\tilde A_{x', s, a, t} = \sum_{x \sim x', x = x'} A_{x, x', s, a, t}$

## Awareness of HIV status

* The proportion of HIV positive adults aware of their HIV status is $v_{x, s, a, t}$
* The proportion of untreated HIV positive adults aware of their HIV status is $\tilde \nu_{x, s, a, t}$ such that
$$
v_{x, s, a, t} = \alpha_{x, s, a, t} + \frac{\tilde v_{R_x, s, a, t}}{1 - \alpha_{x, s, a, t}}
$$

# Likelihood specification

## Notation

* Use $\{x\}$ to denote a collection of districts, $\{s\}$ a collection of sexes and $\{a\}$ a collection of ages for which an outcome is observed

## Household survey data

* Let $\nu$ be a household survey occuring at $T_1$
* Furnishes weighted observations for HIV prevalence $\hat \rho_{\{x\}, \{s\}, \{a\}, \nu}$, ART coverage $\hat \alpha_{\{x\}, \{s\}, \{a\}, \nu}$ and proportion recently infected $\hat \kappa_{\{x\}, \{s\}, \{a\}, \nu}$ with respective Kish effective sample sizes $M^{\hat \theta}_{\{x\}, \{s\}, \{a\}, \nu}$ for $\theta \in \{\rho, \alpha, \kappa\}$
* The observed cases are $Y^{\hat \theta}_{\{x\}, \{s\}, \{a\}, \nu} = M^{\hat \theta}_{\{x\}, \{s\}, \{a\}, \nu} \cdot \hat \theta_{\{x\}, \{s\}, \{a\}, \nu}$
* For HIV prevalence we use a binomial working likelihood
\begin{align}
Y^{\hat \rho}_{\{x\}, \{s\}, \{a\}, \nu} &\sim \text{Bin}(M^{\hat \rho}_{\{x\}, \{s\}, \{a\}, \nu}, \rho_{\{x\}, \{s\}, \{a\}, {T_1}}) \\
\rho_{\{x\}, \{s\}, \{a\}, t} &= \frac{\sum_{x \in \{x\}} \sum_{s \in \{s\}} \sum_{a \in \{a\}} N_{x, s, a, t} \cdot \rho_{x, s, a, t}}{\sum_{x \in \{x\}} \sum_{s \in \{s\}} \sum_{a \in \{a\}} N_{x, s, a, t}}
\end{align}
* For ART coverage we also use a binomial working likelihood
\begin{align}
Y^{\hat \alpha}_{\{x\}, \{s\}, \{a\}, \nu} &\sim \text{Bin}(M^{\hat \alpha}_{\{x\}, \{s\}, \{a\}, \nu}, \alpha_{\{x\}, \{s\}, \{a\}, {T_1}}) \\
\alpha_{\{x\}, \{s\}, \{a\}, t} &= \frac{\sum_{x \in \{x\}} \sum_{s \in \{s\}} \sum_{a \in \{a\}} N_{x, s, a, t} \cdot \rho_{x, s, a, t} \cdot \alpha_{x, s, a, t}}{\sum_{x \in \{x\}} \sum_{s \in \{s\}} \sum_{a \in \{a\}} N_{x, s, a, t} \cdot \rho_{x, s, a, t}}
\end{align}
* For recent infections we also use a binomial working likelihood
\begin{align}
Y^{\hat \kappa}_{\{x\}, \{s\}, \{a\}, \nu} &\sim \text{Bin}(M^{\hat \kappa}_{\{x\}, \{s\}, \{a\}, \nu}, \kappa_{\{x\}, \{s\}, \{a\}, {T_1}}) \\
\kappa_{x, s, a, t} &= 1 - \exp(- \lambda_{x, s, a, t} \cdot \frac{1 - \rho_{x, s, a, t}}{\rho_{x, s, a, t}} \cdot (\Omega_T - \beta_T ) - \beta_T ) \\
\Omega_T &\sim \mathcal{N}(\Omega_{T_0}, \sigma^{\Omega_T}) \\
\beta_T &\sim \mathcal{N}(\beta_{T_0}, \sigma^{\beta_T})
\end{align}

# Clarification notes to discuss with Jeff

* Not sure what $R_x$ means. Is it like $R_x: \mathcal{X} \to \mathcal{X}$
* Some confusion on how to offset terms work -- they seem cool though. Might prefer they not be called $\eta$ which is in my head linear predictor
  * $\eta^\lambda_{R_x, s, a, t}$ aren't these $R_x, s, a, t$ specific rather than 
* Aggregation over age 15-49 uses sum from $a = 15$ to $a = 45$? Perhaps this is because $a = 45$ actually corresponds to the age band 45-49. Edit: yes, $a = 45$ corresponds to 45-49
* Sometimes in subscript we have $x, s, a, t$ and sometimes e.g. $x, s, at$ -- why?
* Why $\sigma^\lambda \sim \mathcal{N}^+(0, 1)$ rather than e.g. $\sigma^\lambda \sim \mathcal{N}^+(0, 2.5)$
* Inconsistent use of $T2$ and $T_2$ in the appendix (I prefer the subscript so try to use throughout here)
* In softmax function, confused as to what this equation is saying: shouldn't there be an equals sign in there before the large brackets?