### 2022-12-18 src/plot-tikz_algorithm-flowchart/algorithm-flowchart.tex
Update flowchart

\documentclass[tikz]{standalone}

% Things to add / improve:
% Try to reduce whitespace a bit
% Arrow goes under rather than through box
% Dashed box around x with i = 1, ..., N
% Dashed box around theta with j = 1, ..., m
% Make it clear AGHQ nodes and weights are being reused
% AGHQ nodes from Gaussian marginal
% More detail than "Marginalise"
% (Maybe) put arrow for normalising constant on the top with arrow that says normalise
% (Maybe) add details about Laplace approximation e.g. making use of starting values, notation for mode and Hessian

\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning}
\usepackage{bm}

\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\z}{\mathbf{z}}
\newcommand{\btheta}{\bm{\theta}}

\begin{document}

\tikzstyle{question} = [rectangle, draw, text width=10em, text centered, rounded corners, minimum height=4em]
\tikzstyle{category} = [rectangle, draw, text width=10em, text centered, rounded corners, minimum height=4em]
\tikzstyle{container} = [rectangle, draw, rounded corners, dashed, inner sep=2em]
\tikzstyle{line} = [draw, -latex']

\begin{tikzpicture}[node distance = 2cm, auto]

    \node[category] at (0, 0) (1) {$p(\theta, \x, \y)$};

    % x
    \node[category] at (0, -3) (a) {$\tilde p_{\text{LA}}(x_i, \btheta, \y)$};
    \node[category] at (0, -6) (b) {$\tilde p_{\text{AQ}}(x_i, \y)$};
    \node[category] at (0, -9) (c) {$\tilde p_{\text{AQ}}(x_i \, | \, \y)$};
    \node[category] at (0, -12) (d) {Image of a nice curve};
    \path[line] (1) -- node[align=center] {Laplace approximation\\Integrate out $\tilde p_{\text{G}}(\x_{-i} \, | \, x_i, \btheta, \y)$}(a);
    \path[line] (a) -- node[align=center] {Integrate out $\theta$\\$\sum_\z \tilde p_{\text{LA}}(x_i, \btheta(\z), \y) \omega(\z)$}(b);
    \path[line] (b) -- node[align=center] {Normalise}(c);
    \path[line] (c) -- node[align=center] {Evaluate $k = 7$ AGHQ nodes\\Use spline interpolation}(d);

    % theta
    \node[category] at (8, 0) (2) {$\tilde p_{\text{LA}}(\btheta, \y)$};
    \node[category] at (8, -3) (3) {$\tilde p_{\text{AQ}}(\y)$};
    \node[category] at (8, -6) (4) {$\tilde p_{\text{AQ}}(\btheta \, | \, \y)$};
    \node[category] at (8, -9) (5) {$\tilde p_{\text{AQ}}(\theta_j \, | \, \y)$};
    \node[category] at (8, -12) (6) {Image of a nice curve};
    \path[line] (1) -- node[align=center] {Laplace approximation\\Integrate out $\tilde p_{\text{G}}(\x \, | \, \btheta, \y)$}(2);
    \path[line] (2) -- node[align=center] {Integrate out $\btheta$ with AGHQ\\Nodes $\z \in \mathcal{Q}(m, k)$\\Adapted $\btheta(\z) = \hat{\btheta} + Lz$\\Weights $\omega(\z):\mathcal{Q} \to \mathbb{R}$}(3);
    \path[line] (3) -- node[align=center] {Normalise} (4);
    \path[line] (2) edge[bend left] node [left] {} (4);
    \path[line] (3) -- (c);
    \path[line] (4) -- node[align=center] {Marginalise} (5);
    \path[line] (5) -- node[align=center] {Evaluate $k = 7$ AGHQ nodes\\Use spline interpolation}(6);

\end{tikzpicture}

\end{document}