### 2023-01-15 src/naomi-simple_model-checks/model-checks.Rmd
Template for model checking notebook

---
title: "Model checking for simplified Naomi"
author:
- name: Adam Howes
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Background**  We have run the simplified Naomi model using a range of inference methods.

  **Task** We check the appropriateness of the model for the data.
---

```{r}
#' Yes!
```