### 2023-04-14 src/docs_18-04-23-lab-group/18-04-23-lab-group.Rmd
Almost finished with slides for group meeting #25

---
title: "Deterministic Bayesian inference methods for the Naomi model"
author: "Adam Howes"
date: "April 2023"
output:
  beamer_presentation:
    latex_engine: pdflatex
    highlight: haddock
    fig_width: 7
    fig_height: 3
    includes:
      in_header: preamble.tex
subtitle: HIV Inference Lab Group Meeting
bibliography: citations.bib
institute: Imperial College London
---
# The Naomi model

:::::::::::::: {.columns}

::: {.column width=.65}

* Naomi is a complicated spatio-temporal evidence synthesis model
* Used by countries to produce HIV estimates in a yearly process supported by UNAIDS
* Fast inference is important to allow for interactive review and development of estimates
* Inference for Naomi is currently conducted using Template Model Builder (`TMB`) [@kristensen2015tmb]

:::

::: {.column width=.35}

```{r, echo=FALSE, fig.cap="A supermodel", out.width = '65%'}
knitr::include_graphics("naomi_hex.png")
```

:::

::::::::::::::

#

```{r, echo=FALSE, fig.cap="Example of the user interface from \\texttt{https://naomi.unaids.org/}", out.width = '70%'}
knitr::include_graphics("naomi_user.png")
```

# Why do we use `TMB`

1. It runs quickly
2. It is flexible enough to be compatible with the model
3. We don't have better viable options

# What problem we are trying to solve

* Ideally we want exact Bayesian inference: compute the posterior distribution of the parameters of the model given the data
* Computationally this amounts to solving a difficult integral
* We can't solve this, but we can give approximate answers

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]

Goal: approximate this integral for Naomi better than \texttt{TMB} does, without taking too long. In doing so, more accurately reflect uncertainty over hyperparameters

\end{tcolorbox}
\end{center}

# Two deterministic methods

* We use two deterministic^[In contrast to the most famous approximate Bayesian inference method, Markov chain Monte Carlo, which is fundamentally stochastic.] methods to approximate our integral

1. The Laplace approximation
2. Quadrature

# The Laplace approximation

* If you pretend the posterior distribution is a Gaussian, computation is easy

#

```{r message=FALSE, echo=FALSE,  fig.cap="A Gamma prior with $a = 3$ and $b = 1$."}
cbpalette <- c("#56B4E9","#009E73", "#E69F00", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

a <- 3
b <- 1

prior <- ggplot(data = data.frame(x = c(0, 10)), aes(x)) +
  stat_function(fun = dgamma, n = 500, args = list(shape = a, rate = b), col = cbpalette[1]) +
  annotate("text", x = 6, y = 0.15, label = "Gamma(3, 1)", col = cbpalette[1], size = 5) +
  labs(x = "", y = "") +
  theme_minimal()

prior
```

#

```{r message=FALSE, echo=FALSE,  fig.cap="Draw 3 points from $\\text{Poisson}(3)$, then compute the posterior."}
set.seed(2)
y <- rpois(3, lambda = 2)

posterior <- prior +
  geom_point(data = data.frame(x = y, y = 0), aes(x = x, y = y), inherit.aes = FALSE, alpha = 0.7, size = 2) +
  stat_function(data = data.frame(x = c(0, 10)), aes(x), fun = dgamma, n = 500, args = list(shape = a + sum(y), rate = b + length(y)), col = cbpalette[2]) +
  annotate("text", x = 5, y = 0.25, label = "Gamma(9, 4)", col = cbpalette[2], size = 5)

posterior
```

#

```{r message=FALSE}
fn <- function(x) dgamma(x, a + sum(y), b + length(y), log = TRUE)

# Here we are using numerical derivatives
ff <- list(
  fn = fn,
  gr = function(x) numDeriv::grad(fn, x),
  he = function(x) numDeriv::hessian(fn, x)
)

opt_bfgs <- aghq::optimize_theta(
  ff, 1, control = aghq::default_control(method = "BFGS")
)
```

# Laplace approximation

```{r}
laplace <- posterior +
  stat_function(
    data = data.frame(x = c(0, 10)),
    aes(x),
    fun = dnorm,
    n = 500,
    args = list(mean = opt_bfgs$mode, sd = sqrt(1 / opt_bfgs$hessian)),
    col = cbpalette[3]
  )
```

# Laplace approximation

```{r message=FALSE, echo=FALSE,  fig.cap="The Laplace approximation in this case is good near the mode but not in the tails."}
laplace +
  annotate("text", x = 5.5, y = 0.35, label = "Laplace approximation", col = cbpalette[3], size = 5)
```

# Computation of the Laplace approximation

* This computation was simple, and involved

1. Optimising a function
2. Taking the mode and the Hessian at the mode

# The marginal Laplace approximation

* If we don't want to pretend the whole posterior distribution is Gaussian, another option is to pretend some of its marginals are
* This is how `TMB` works: it's up to the user to choose which parameters should be Gaussian using the `random` option

# Which parameters should we treat as Gaussian?

* In spatio-temporal statistics we have data indexed by space and time
* We use random effects also indexed by space and time to model this data
* Spatio-temporal fields can be big
* Willing to make assumptions about how things vary over spacetime

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]

$\implies$ treat the latent field parameters as Gaussian! That's the majority of the integral done.

\end{tcolorbox}
\end{center}

# What about the hyperparameters?

* `TMB` uses optimisation to find the hyperparameters which maximise the marginal Laplace approximation
* This is the "outer" optimisation loop, where the "inner" is for computation of the Gaussian distribution

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]

Inference for the latent field is based on a single value of the hyperparameters (the mode) -- so called empirical Bayes $\implies$ no uncertainty in the hyperparameters taken into account! How can you sleep at night.

\end{tcolorbox}
\end{center}

# Quadrature

* This brings us to our other method for solving integrals deterministically
* Say we have a function, then quadrature has two ingredients

1. Nodes: points to evaluate the function at
2. Weights: importance of function evaluation at that point

* You evaluate the function at the nodes, then do a weighted sum to calculate your integral

# Trapezoid rule example

* Let's compute $\int_0^\pi \sin(x) \text{d}x = 2$ using quadrature

```{r}
trapezoid_rule <- function(x, spacing) {
  # Assumes nodes are evenly spaced
  w <- rep(spacing, length(x)) # Weights given by space between nodes
  w[1] <- w[1] / 2 # Apart from the first which is halved
  w[length(x)] <- w[length(x)] / 2 # And the last, also halved
  sum(w * x) # Compute the weighted sum
}
```

#

```{r message=FALSE, echo=FALSE,  fig.cap="With 10 nodes it's 0.02 off."}
finegrid <- seq(0, pi, length.out = 1000)

plot <- data.frame(x = finegrid, y = sin(finegrid)) %>%
  ggplot(aes(x = x, y = y)) +
   geom_line(col = cbpalette[1]) +
   theme_minimal() +
   labs(x = "x", y = "sin(x)")

trapezoid_plot <- function(N) {
  grid <- seq(0, pi, length.out = N)
  int <- trapezoid_rule(x = sin(grid), spacing = grid[2] - grid[1])

  plot + 
    geom_bar(
      data = data.frame(x = grid, y = sin(grid)),
      aes(x = x, y = y), alpha = 0.7, stat = "identity",
      inherit.aes = FALSE, fill = cbpalette[2]) +
    theme_minimal() +
    labs(
      subtitle = paste0("Number of nodes: ", N, "\nTrapezoid rule estimate: ", round(int, 3), "\nTruth: 2"),
      x = "x", y = "sin(x)"
    )
}

trapezoid_plot(N = 10)
```

#

```{r message=FALSE, echo=FALSE, fig.cap="With 30 nodes it's 0.002 off."}
trapezoid_plot(N = 30)
```

#

```{r message=FALSE, echo=FALSE, fig.cap="With 100 nodes it's pretty much correct."}
trapezoid_plot(N = 100)
```

# Adaptive Gauss-Hermite quadrature

* Gauss-Hermite quadrature is a method for picking nodes and weights based on the theory of polynomial interpolation
* It works especially well for statistical problems where the integrand looks like something multiplied by a Gaussian distribution
* The "adaptive" part means the nodes and weights are changed depending on the integrand -- this makes sense, especially when the integrand is also a function of the data
* Implemented by the `aghq` package [@stringer2021implementing]

#

```{r, echo=FALSE, message=FALSE, fig.cap="Unadapted points in two dimensions with $k = 3$."}
mu <- c(1, 1.5)
cov <- matrix(c(2, 1, 1, 1), ncol = 2)

obj <- function(theta) {
  mvtnorm::dmvnorm(theta, mean = mu, sigma = cov)
}

grid <- expand.grid(
  theta1 = seq(-2, 5, length.out = 700),
  theta2 = seq(-2, 5, length.out = 700)
)

ground_truth <- cbind(grid, pdf = obj(grid))

plot0 <- ggplot(ground_truth, aes(x = theta1, y = theta2, z = pdf)) +
  geom_contour(col = cbpalette[1]) +
  coord_fixed(xlim = c(-2, 4.5), ylim = c(-2, 4.5), ratio = 1) +
  labs(x = "", y = "") +
  theme_minimal()

gg <- mvQuad::createNIGrid(2, "GHe", 3)

add_points <- function(plot0, gg) {
  plot0 +
    geom_point(
      data = mvQuad::getNodes(gg) %>%
              as.data.frame() %>%
              mutate(weights = mvQuad::getWeights(gg)), 
      aes(x = V1, y = V2, size = weights),
      alpha = 0.8,
      col = cbpalette[2],
      inherit.aes = FALSE
    ) +
    scale_size_continuous(range = c(1, 2))
}

add_points(plot0, gg) +
  labs(size = "Weight", caption = "")
```

#

```{r, echo=FALSE, message=FALSE, fig.cap="Add the mean $z + \\hat \\theta$."}
gg2 <- gg
mvQuad::rescale(gg2, m = mu, C = diag(c(1, 1)), dec.type = 2)

add_points(plot0, gg2) +
  labs(size = "Weight", caption = "")
```

#

```{r, echo=FALSE, message=FALSE, fig.cap="First option: rotate by the lower Cholesky $Lz + \\hat \\theta$."}
gg3 <- gg
mvQuad::rescale(gg3, m = mu, C = cov, dec.type = 2)

add_points(plot0, gg3) +
  labs(size = "Weight", caption = "")
```

#

```{r, echo=FALSE, message=FALSE, fig.cap="Second option: rotate using the eigendecomposition $E \\Lambda^{1/2} z + \\hat \\theta$."}
gg3 <- gg
mvQuad::rescale(gg3, m = mu, C = cov, dec.type = 1)

add_points(plot0, gg3) +
  labs(size = "Weight", caption = "")
```

# The plan?

:::::::::::::: {.columns}

::: {.column width=.55}

* Write and implement an algorithm for fast approximate Bayesian inference using the Laplace approximation and quadrature for the Naomi model
* Use `TMB` for writing the model in C++ and implementing the Laplace approximation via automatic differentitation
* Use adaptive Gauss-Hermite quadrature to integrate the hyperparameters

:::

::: {.column width=.45}

```{r, echo=FALSE, fig.cap="One of the best engineering schools in Canada", out.width = '90%'}
knitr::include_graphics("the-plan.png")
```

:::

::::::::::::::

# Long prophecised

> My main comment is that several aspects of the computational machineery that is presented by Rue and his colleagues **could benefit from the use of a numerical technique known as automatic differentiation (AD)** ... By the use of AD one could obtain a system that is automatic from a user's perspective... the benefit would be a fast, flexible and easy-to-use system for doing Bayesian analysis in models with Gaussian latent variables

-- Hans J. Skaug (coauthor of `TMB`), RSS discussion of @rue2009approximate

# One challenge

* For Malawi, Naomi has 24 hyperparameters: too many for a dense grid
* Proposed solution use principal components analysis and keep only first $s < 24$ dimensions
* We will use $k = 3$ points per dimension

#

```{r, echo=FALSE, message=FALSE, fig.cap="Illustration of what this looks like for the 2D example, keeping only the first principal component, $s = 1$."}
gg3 <- gg
mvQuad::rescale(gg3, m = mu, C = cov, dec.type = 2)

add_points(plot0, gg3) +
  labs(size = "Weight", caption = "")
```

#

```{r, echo=FALSE, fig.cap="This Scree plot suggests 10 or so dimensions is enough. We use $s = 8$.", out.width = '80%'}
knitr::include_graphics("depends/tv-plot.png")
```

#

```{r, echo=FALSE, fig.cap="With 8 dimenions, the covariance matrix is accurately reproduced.", out.width = '80%'}
knitr::include_graphics("depends/reduced-rank-plot.png")
```

# Comparison of posterior distributions

* Use test data from Malawi
* Run HMC for as long as it takes to get a good answer
* Compare output of `TMB` with `aghq`: which is more similar to HMC?
* There are many ways you might do this: I'll show you Kolmogorov-Smirnov test results

#

Worse case TMB plot demonstrating KS test

#

Overall KS results plot

# Conclusions

# Thanks for listening!

* Working on a paper "Fast approximate Bayesian inference for small-area estimation of HIV indicators using the Naomi model" based on this work
  * Joint with Alex Stringer (Waterloo), Seth Flaxman (Oxford), Jeff Eaton (Imperial)
* Code and notebooks for this project are at `athowes.github.io/elgm-inf`

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]

Let me know if you'd be up for being an early reader!

\end{tcolorbox}
\end{center}

# Risk group retrospective

```{r, echo=FALSE, fig.cap="Available at \\texttt{athowes.github.io}", out.width = '90%'}
knitr::include_graphics("blog-post.png")
```

# References {.allowframebreaks}