### 2023-05-09 src/naomi-simple_mmd/mmd.Rmd
Calculate MMD with kernlab #42

---
title: "Inference methods comparison for the simplified Naomi model using maximum mean discrepancy"
author:
- name: Adam Howes
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Background**  We have run the simplified Naomi model using a range of inference methods: `TMB`, `aghq` and `tmbstan`.

  **Task** In this report, we compare the accuracy of the posterior distributions obtained from these inference methods using maximum mean discrepancy.
---

# Background

We compare the inference results from `TMB`, `aghq`, `adam`, and `tmbstan`.
Import these inference results as follows:

```{r}
tmb <- readRDS("depends/tmb.rds")
aghq <- readRDS("depends/aghq.rds")
tmbstan <- readRDS("depends/tmbstan.rds")

depends <- yaml::read_yaml("orderly.yml")$depends
```

Check that the parameters (latent field, hyperparameters, model outputs) sampled from each of the four methods are the same:

```{r}
stopifnot(names(tmb$fit$sample) == names(aghq$quad$sample))
stopifnot(names(tmb$fit$sample) == names(tmbstan$mcmc$sample))
```

# Maximum mean discrepancy

```{r}
mmd1 <- kernlab::kmmd(matrix(tmb$fit$sample$beta_rho), matrix(tmbstan$mcmc$sample$beta_rho))
mmd2 <- kernlab::kmmd(matrix(aghq$quad$sample$beta_rho), matrix(tmbstan$mcmc$sample$beta_rho))

mmd1
mmd2
```