### 2023-05-10 src/naomi-simple_exceedance/exceedance.Rmd
Short notes on excursions #35

---
title: "Exceedance probability comparison for the simplified Naomi model"
author:
- name: Adam Howes
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
bibliography: citations.bib
abstract: |
  **Background**  We have run the simplified Naomi model using a range of inference methods: `TMB`, `aghq`, `adam` and `tmbstan`.

  **Task** In this report, we compare the accuracy of the posterior distributions obtained from these inference methods using exceedance probabilities.
---

# Background

We compare the inference results from `TMB`, `aghq`, `adam`, and `tmbstan`.
Import these inference results as follows:

```{r}
tmb <- readRDS("depends/tmb.rds")
aghq <- readRDS("depends/aghq.rds")
tmbstan <- readRDS("depends/tmbstan.rds")
```

# Second 90

The 90-90-90 treatment targets are that

* 90% of PLHIV know their status
* 90% of PLHIV who kno wtheir status are on antiretroviral therapy (ART)
* 90% of PLHIV on ART have suppressed viral load

To meet the second 90, we require that $90\% \times 90\% = 81\%$ of PLHIV are on ART.

```{r}
mf_out <- tmb$naomi_data$mf_out

mf_out_fine <- mf_out %>%
  tibble::rownames_to_column("id") %>%
  mutate(id = as.numeric(id)) %>%
  filter(
    area_id %in% paste0("MWI_4_", 1:32, "_demo"),
    sex %in% c("male", "female"),
    age_group %in%
      c(
        "Y000_004", "Y005_009", "Y010_014", "Y015_019", "Y020_024", "Y025_029",
        "Y025_034", "Y030_034", "Y035_039", "Y040_044", "Y045_049", "Y050_054",
        "Y055_059", "Y060_064", "Y065_069", "Y070_074", "Y075_079", "Y080_999"
      )
  )

tmbstan_second90 <- apply(tmbstan$mcmc$sample$alpha_t1_out[mf_out_fine$id, ], 1, function(x) sum(x > 0.81) / length(x))
tmb_second90 <- apply(tmb$fit$sample$alpha_t1_out[mf_out_fine$id, ], 1, function(x) sum(x > 0.81) / length(x))
aghq_second90 <- apply(aghq$quad$sample$alpha_t1_out[mf_out_fine$id, ], 1, function(x) sum(x > 0.81) / length(x))

plot(tmbstan_second90, tmb_second90)
plot(tmbstan_second90, aghq_second90)

cor(tmbstan_second90, tmb_second90)
cor(tmbstan_second90, aghq_second90)

mse_tmb_second90 <- mean((tmbstan_second90 - tmb_second90)^2)
mse_aghq_second90 <- mean((tmbstan_second90 - aghq_second90)^2)
round(100 * (mse_aghq_second90 - mse_tmb_second90) / mse_tmb_second90)

rmse_tmb_second90 <- sqrt(mse_tmb_second90)
rmse_aghq_second90 <- sqrt(mse_aghq_second90)
round(100 * (rmse_aghq_second90 - rmse_tmb_second90) / rmse_tmb_second90)
```

# High incidence strata

Above 1% incidence is considered high, and could be met with intensified interventions.

* Very high (>3%)
* High (1-3%)
* Moderate (0.3-1%)
* Low (<0.3%)

```{r}
tmbstan_1inc <- apply(tmbstan$mcmc$sample$lambda_t1_out[mf_out_fine$id, ], 1, function(x) sum(x > 0.01) / length(x))
tmb_1inc <- apply(tmb$fit$sample$lambda_t1_out[mf_out_fine$id, ], 1, function(x) sum(x > 0.01) / length(x))
aghq_1inc <- apply(aghq$quad$sample$lambda_t1_out[mf_out_fine$id, ], 1, function(x) sum(x > 0.01) / length(x))

plot(tmbstan_1inc, tmb_1inc)
plot(tmbstan_1inc, aghq_1inc)

cor(tmbstan_1inc, tmb_1inc)
cor(tmbstan_1inc, aghq_1inc)

mse_tmb_1inc <- mean((tmbstan_1inc - tmb_1inc)^2)
mse_aghq_1inc <- mean((tmbstan_1inc - aghq_1inc)^2)
round(100 * (mse_aghq_1inc - mse_tmb_1inc) / mse_tmb_1inc)

rmse_tmb_1inc <- sqrt(mse_tmb_1inc)
rmse_aghq_1inc <- sqrt(mse_aghq_1inc)
round(100 * (rmse_aghq_1inc - rmse_tmb_1inc) / rmse_tmb_1inc)
```

# Unmet treatment need

Number of people needing treatment is a function of ART coverage, HIV prevalence and population size.
What summaries of unmet treatment need might be important to calculate?

```{r}
#' To-do!
```

# Excursion and contour uncertainty regions

Based on @bolin2015excursion.

* $D_m = \{s: P(x(s) > u) \geq 1 - \alpha\}$ runs into the problem of multiple hypothesis testing
* Care about $P(x(s) > u, s \in D_m)$ which is smaller than $1 - \alpha$
* You can use the marginals and adjust the threshold to control type 1 error, false discovery rate, or posteior probabilites
* Bolin and Lindgren use sequential importance sampling to estimate the joint probabilities
* $x(s) > u$ is called a positive excursion set, and $x(s) < u$ is called a negative excursion set
* The `excursions` package can calculate excursion sets, contour credible regions and level avoiding sets, excursion functions, contour maps and their quality measures, and simultaneous confidence bands