### 2023-05-23 src/docs_bioinference-poster/bioinference-poster.Rmd
Update template for BioInference poster #36

---
poster_height: "1189mm"
poster_width: "841mm"
font_family: "Open Sans"
# Essentials
title: Fast approximate Bayesian inference for small-area estimation of HIV indicators using the Naomi model
# titlebox_height: "10%"
author:
  - name: "Adam Howes"
    affil: 1, 2
    main: true
    email: ath19@ic.ac.uk
  - name: "Alex Stringer"
    affil: 3
  - name: "Seth R. Flaxman"
    affil: 4
  - name: "Jeffrey W. Eaton"
    affil: 5, 2
affiliation:
  - num: 1
    address: Department of Mathematics, Imperial College London
  - num: 2
    address: MRC Centre for Global Infectious Disease Analysis, School of Public Health, Imperial College London
  - num: 3
    address: Department of Statistics and Actuarial Science, University of Waterloo
  - num: 4
    address: Department of Computer Science, University of Oxford
  - num: 5
    address: Center for Communicable Disease Dynamics, Department of Epidemiology, Harvard T.H. Chan School of Public Health
logoleft_name: "bioinference-joined.png"
# Style & formatting
column_numbers: 3
title_textsize: "75px"
author_textsize: "55px"
authorextra_textsize: "35px"
affiliation_textsize: "35px"
caption_fontsize: "20px"
# Title box
title_textcol: "#000000"
author_textcol: "#000000"
affiliation_textcol: "#000000"
# Main
main_fontfamily: "Open Sans"
# Poster body options
body_bgcol: "#ffffff"
body_textsize: "36px"
body_textcol: "#000000"
reference_textsize: "25px"
sectitle_textsize: "40px"
link_col: "#000000"
# Standard options
bibliography: citations.bib
output: 
  posterdown::posterdown_html:
    self_contained: false
    pandoc_args: --mathjax
    highlight: espresso
    number_sections: false
---

```{css, echo=FALSE}
div.logo_left{
  width: 35%;
  background-color: #E5E5E5;
}
div.poster_title{
  width: 60%;
}
div.title_container{
  background-color: #E5E5E5;
}
code{
  background-color: #ffffff;
  color: #000000;
}
.poster_body{
  column-rule-color: #ffffff;
}
```

## Summary

* Approximate Bayesian inference method using Laplace approximation, adaptive Gauss-Hermite quadrature and principal component analysis
* Motivated by an evidence synthesis model for small-area estimation of HIV indicators in sub-Saharan Africa
* Implemented as a part of the `aghq` package [@stringer2021implementing], allowing flexible use of the method for any model with a `TMB` C++ user template

## The Naomi HIV model

* District-level model of HIV indicators [@eaton2021naomi] which synthesises data from household surveys, antenatal care (ANC) clinics, and routine service provision of antiretroviral therapy (ART)
  * Combining evidence from multiple data sources helps overcome the limitations of any one
  * Small-area estimation methods to overcome small district-level sample sizes
* Yearly estimation process: model run interactively by country teams using a web-app `naomi.unaids.org`
  * Figure \@ref(fig:fig1) illustrates the seven stages of using the app
* Inference conducted in minutes using empirical Bayes and a Gaussian approximation via Template Model Builder `TMB` [@kristensen2016tmb]
* It would take days to get accurate answers with MCMC via `tmbstan` [@monnahan2018no], and this is not practical in this setting
* Motivates looking for a fast, approximate approach, that takes uncertainty in hyperparameters into account

$$\\[0.1in]$$

```{r fig1, echo=FALSE, fig.align='center', out.width='100%', fig.cap='Model fitting occurs interactively in stages.'}
knitr::include_graphics("fig1.png")
```

$$\\[0.1in]$$

## Extended latent Gaussian models

* Latent Gaussian models (LGMs) [@rue2009approximate] are three stage hierarchical models with observations $y$, Gaussian latent field $x$ and hyperparameters $\theta$
* In an LGM the conditional mean depends on exactly one structured additive predictor $\mu_i = g(\eta_i)$ with $g: \mathbb{R} \to \mathbb{R}$
  * The `R-INLA` implementation of integrated nested Laplace approximations applies only to LGMs, because ELGM precision matrices are not as sparse
* Extended latent Gaussian models (ELGM) remove this requirement such that $\mu_i = g(\eta_{\mathcal{J_i}})$ where $g_i: \mathbb{R}^{|\mathcal{J_i}|} \to \mathbb{R}$ and $\mathcal{J_i}$ is some set of indices
  * Allows a higher degree of non-linearity in the model
* Naomi is an ELGM, not an LGM, because it includes complex dependency structures:
  1. Incidence depends on prevalence and ART coverage
  2. Incidence ane prevalence linked to recent infection
  3. ANC offset from household survey
  4. ART coverage and recent infection are products
  5. Observed data are aggregated finer processes
  6. ART attendance uses the multinomial
  7. Multiple link functions
* We extend work of @stringer2022fast in this setting to the challenging Naomi ELGM
* Though we focus on Naomi here, the HIV Inference Group (`hiv-inference.org`) works on many other complex models, challenging for existing Bayesian inference methods, which require flexible modelling tools

## Inference procedure

* **Laplace approximation** Integrate out variables using a Gaussian approximation to the denominator
$$
p(\theta, y) \approx \tilde p_\texttt{LA}(\theta, y) = \frac{p(y, x, \theta)}{\tilde p_\texttt{G}(x \, | \, \theta, y)} \Big\rvert_{x = \hat x(\theta)}
$$
where $\tilde p_\texttt{G}(x \, | \, \theta, y) = \mathcal{N}(x \, | \, \hat x(\theta), \mathbf{H}(\theta)^{-1})$
  * Use automatic differentiation via `CppAD` in `TMB`
* **Adaptive Gauss-Hermite Quadrature**
$$
\int_\Theta p(\theta) \text{d} \theta \approx \lvert L \rvert \sum_{z \in \mathcal{Q}(m, k)} p(\hat \theta + Lz) \omega(z)
$$
where the Gauss-Hermite quadrature rule $z \in \mathcal{Q}(m, k)$ with $m = \dim(\theta)$ and $k$ points per dimension is adapted based upon
  * The mode $\hat \theta = \text{argmax}_{\theta \in \Theta} p(\theta)$
  * A matrix decomposition $LL^\top = - \partial_\theta^2 \log p(\theta) \rvert_{\theta = \hat \theta}$
* Keep first $s < m$ **principal components** to get size $k^s$ grid

$$\\[0.1in]$$

## Application to Malawi data

```{r echo=FALSE, message=FALSE, warning=FALSE}
time_taken <- readr::read_csv("depends/time_taken.csv")
```

```{r fig3, echo=FALSE, fig.align='center', out.width='100%', fig.cap='District-level model outputs for adults 15-49 in January 2016. Adapted from Eaton et al. 2021.'}
knitr::include_graphics("fig3.png")
```

* Relatively small country but still a large model: latent field $\text{dim}(x) = 491$, hyperparameters $\text{dim}(\theta) = 24$
* Fit three inference methods (using one C++ template):
  * TMB (`r round(time_taken$TMB, 0)` mins)
  * PCA-AGHQ (`r round(time_taken$aghq, 0)` mins): $k = 3, s = 8$
  * NUTS (`r round(time_taken$tmbstan, 1)` days): 4 chains of 100,000 thinned by 40 (required for good diagnostics)
* Figure \@ref(fig:fig3) illustrates example model outputs: HIV prevalence, ART coverage, HIV incidence, and number of new infections, at the district level
* Compare hyperparameter, latent field, and output posteriors based on Kolmogorov-Smirnov tests, Pareto-smoothed importance sampling, and maximum mean discrepancy

## Future directions

* Laplace marginals with matrix algebra approximations [@wood2020simplified] to speed up latent field marginal calculations
* Further methods for allocation of effort to important dimensions

**Interested?** Working notebooks and R code available from `github.com/athowes/elgm-inf`. Or get in touch:

* `r fontawesome::fa("home", fill = "black")` `athowes.github.io`
* `r fontawesome::fa("envelope", fill = "black")` `ath19@ic.ac.uk`
* `r fontawesome::fa("twitter", fill = "black")` `adamhowes`

**Funding** AH was supported by the EPSRC and Bill & Melinda Gates Foundation. This research was supported by the MRC Centre for Global Infectious Disease Analysis.

## References

<font size="5"> <div id="refs" class="references"></div> </font>