### 2023-06-05 src/docs_bioinference-poster/bioinference-poster.Rmd
Just need to add second 90 plot then finished with BioInference poster #36

---
poster_height: "1189mm"
poster_width: "841mm"
font_family: "Open Sans"
# Essentials
title: Fast approximate Bayesian inference for small-area estimation of HIV indicators using the Naomi model
# titlebox_height: "10%"
author:
  - name: "Adam Howes"
    affil: 1, 2
    main: true
    email: ath19@ic.ac.uk
  - name: "Alex Stringer"
    affil: 3
  - name: "Seth R. Flaxman"
    affil: 4
  - name: "Jeffrey W. Eaton"
    affil: 5, 2
affiliation:
  - num: 1
    address: Department of Mathematics, Imperial College London
  - num: 2
    address: MRC Centre for Global Infectious Disease Analysis, School of Public Health, Imperial College London
  - num: 3
    address: Department of Statistics and Actuarial Science, University of Waterloo
  - num: 4
    address: Department of Computer Science, University of Oxford
  - num: 5
    address: Center for Communicable Disease Dynamics, Department of Epidemiology, Harvard T.H. Chan School of Public Health
logoleft_name: "bioinference-joined.png"
# Style & formatting
column_numbers: 3
title_textsize: "75px"
author_textsize: "55px"
authorextra_textsize: "35px"
affiliation_textsize: "35px"
caption_fontsize: "20px"
# Title box
title_textcol: "#000000"
author_textcol: "#000000"
affiliation_textcol: "#000000"
# Main
main_fontfamily: "Open Sans"
# Poster body options
body_bgcol: "#ffffff"
body_textsize: "36px"
body_textcol: "#000000"
reference_textsize: "25px"
sectitle_textsize: "40px"
link_col: "#000000"
# Standard options
bibliography: citations.bib
output: 
  posterdown::posterdown_html:
    self_contained: false
    pandoc_args: --mathjax
    highlight: espresso
---

```{css, echo=FALSE}
div.logo_left{
  width: 35%;
  background-color: #E5E5E5;
}
div.poster_title{
  width: 60%;
}
div.title_container{
  background-color: #E5E5E5;
}
code{
  background-color: #ffffff;
  color: #000000;
}
.poster_body{
  column-rule-color: #ffffff;
}
```

## Summary

* We developed an approximate Bayesian inference method using Laplace approximation, adaptive Gauss-Hermite quadrature and principal component analysis
* Motivated by an evidence synthesis model for small-area estimation of HIV indicators in sub-Saharan Africa
* Implemented as a part of the `aghq` package [@stringer2021implementing], allowing flexible use of the method for any model with a Template Model Builder `TMB` [@kristensen2016tmb] C++ user template

## 1. The Naomi HIV model

* District-level model of HIV indicators [@eaton2021naomi] which synthesises data from 1) household surveys, 2) antenatal care (ANC) clinics, and 3) routine service provision of antiretroviral therapy (ART)
  * Combining evidence from multiple data sources helps overcome the limitations of any one
  * Small-area estimation methods to overcome small district-level sample sizes
* Yearly estimation process: model run interactively by country teams using a web-app `naomi.unaids.org`
  * Figure \@ref(fig:fig1) illustrates the seven stages of using the app
* Inference conducted in minutes using empirical Bayes and a Gaussian approximation
* It would take days to get accurate answers with MCMC via `tmbstan` [@monnahan2018no], and this is not practical in this setting
* We are looking for a fast, approximate approach, that properly takes uncertainty in hyperparameters into account

$$\\[0.1in]$$

```{r fig1, echo=FALSE, fig.align='center', out.width='100%', fig.cap='Model fitting occurs interactively in stages.'}
knitr::include_graphics("fig1.png")
```

$$\\[0.1in]$$

## 2. Extended latent Gaussian models

* Latent Gaussian models (LGMs) [@rue2009approximate] are three stage hierarchical models with observations $y$, Gaussian latent field $x$ and hyperparameters $\theta$
* In an LGM the conditional mean depends on exactly one structured additive predictor $\mu_i = g(\eta_i)$ with $g: \mathbb{R} \to \mathbb{R}$
* Extended latent Gaussian models (ELGM) remove this requirement such that $\mu_i = g(\eta_{\mathcal{J_i}})$ where $g_i: \mathbb{R}^{|\mathcal{J_i}|} \to \mathbb{R}$ and $\mathcal{J_i}$ is some set of indices
  * Allows a higher degree of non-linearity in the model
* Naomi is an ELGM, not an LGM, because it includes complex dependency structures:
  1. Incidence depends on prevalence and ART coverage
  2. Incidence ane prevalence linked to recent infection
  3. ANC offset from household survey
  4. ART coverage and recent infection are products
  5. Observed data are aggregated finer processes
  6. ART attendance uses the multinomial
  7. Multiple link functions
* We extend work of @stringer2022fast in this setting to the challenging Naomi ELGM
* Though we focus on Naomi, the HIV Inference Group (`hiv-inference.org`) works on many other complex models, challenging for existing Bayesian inference methods, which require flexible modelling tools

## 3. Inference procedure

* **Laplace approximation** Integrate out latent field using a Gaussian approximation to the denominator
$$
p(\theta, y) \approx \tilde p_\texttt{LA}(\theta, y) = \frac{p(y, x, \theta)}{\tilde p_\texttt{G}(x \, | \, \theta, y)} \Big\rvert_{x = \hat x(\theta)},
$$
where $\tilde p_\texttt{G}(x \, | \, \theta, y) = \mathcal{N}(x \, | \, \hat x(\theta), \hat H(\theta)^{-1})$
  * Use automatic differentiation via `CppAD` in `TMB`

```{r aghq, echo=FALSE, fig.align='center', fig.cap='Demonstration of PCA-AGHQ.'}
knitr::include_graphics("fig2.png")
```

* **Adaptive Gauss-Hermite Quadrature** (AGHQ) perform quadrature over the hyperparameters
$$
\int_\Theta p_\texttt{LA}(\theta, y) \text{d} \theta \approx \lvert L \rvert \sum_{z \in \mathcal{Q}(m, k)} p_\texttt{LA}(\hat \theta + Lz, y) \omega(z),
$$
where the Gauss-Hermite quadrature rule $\{z \in \mathcal{Q}(m, k), \omega\}$ with $m = \dim(\theta)$ and $k$ points per dimension is adapted based upon
  * The mode $\hat \theta = \text{argmax}_{\theta \in \Theta} p_\texttt{LA}(\theta, y)$
  * A matrix decomposition $LL^\top = - \partial_\theta^2 \log p_\texttt{LA}(\theta, y) \rvert_{\theta = \hat \theta}$
* Use the spectral decomposition $L = E\Lambda^{1/2}$ and keep only the first $s < m$ **principal components** (PCA-AGHQ)

## 4. Application to Malawi

```{r fig3, echo=FALSE, fig.align='center', fig.cap='District-level model outputs for adults 15-49 in January 2016. Adapted from Eaton et al. 2021.'}
knitr::include_graphics("fig3.png")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
time_taken <- readr::read_csv("depends/time_taken.csv")
```

* Malawi is a relatively small country but still has latent field $\text{dim}(x) = 491$ and hyperparameters $\text{dim}(\theta) = 24$

| Method | Description | Time |
|-----------|----------|------|
| TMB | Baseline | `r round(time_taken$TMB, 0)` secs |
| PCA-AGHQ | Ours | `r round(time_taken$aghq, 0)` hour |
| NUTS | Gold-standard | `r round(time_taken$tmbstan, 1)` days |

* For PCA-AGHQ $k = 3$ and $s = 8$ chosen using Scree plot to explain ~90% of variance
* For NUTS 4 chains of 100,000 thinned by 40 were required for good diagnostics
* Kolmogorov-Smirnov (KS) test based on the maximum difference between marginal ECDFs
  * Reduces average KS distance from 0.08 to 0.075
* Also considering joints via Pareto-smoothed importance sampling and maximum mean discrepancy
* Naomi can be used to assess probabilites targets have been met e.g. 90% of those who know their HIV status are on ART ("second 90")
  * Reduces RMSE for estimating second 90 exceedance probabilities by 10%

```{r}
# knitr::include_graphics("fig4.png")
# Figure of Second 90 here
```

## 5. Future directions

* Laplace marginals with matrix algebra approximations [@wood2020simplified] to speed up calculations
* Further methods for allocation of effort to "important" dimensions of hyperparameter grid

**Interested?** Working notebooks and R code available from `github.com/athowes/elgm-inf`. Or get in touch:

* `r fontawesome::fa("home", fill = "black")` `athowes.github.io`
* `r fontawesome::fa("envelope", fill = "black")` `ath19@ic.ac.uk`
* `r fontawesome::fa("twitter", fill = "black")` `adamhowes`

**Funding** AH was supported by the EPSRC and Bill & Melinda Gates Foundation. This research was supported by the MRC Centre for Global Infectious Disease Analysis.

## References

<font size="5"> <div id="refs" class="references"></div> </font>