### 2023-07-04 src/astro/astro.Rmd
Fixes to astro example -- now understand where we are (needing a method for hyper marginals from PCA-AGHQ and EB) #46

---
title: "Milky way mass estimation using PCA-AGHQ"
author:
- name: Adam Howes
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
bibliography: citations.bib
abstract: |
  **Task** We reanalyse the "estimating the mass of the Milky Way" example from @bilodeau2021stochastic.
---

Originally Alex used the `ipoptr` package.
However, it is now not possible, or at least very difficult, to install it (as of 01/04/2023).
We will here use the `nloptr` package instead.

```{r}
data(gcdatalist, package = "aghq")
precompile()

set.seed(4365789)

globalpath <- normalizePath(tempdir(), winslash = "/")
plotpath <- normalizePath(file.path(globalpath, "astro"), winslash = "/", mustWork = FALSE)
if (!dir.exists(plotpath)) dir.create(plotpath)
```

Get the TMB template from the `aghq` package, the compile and load it:

```{r}
file.copy(normalizePath(system.file("extsrc/01_astro.cpp", package = "aghq"), winslash = "/"), globalpath)
compile(normalizePath(file.path(globalpath, "01_astro.cpp"), winslash = "/"))
dyn.load(normalizePath(dynlib(file.path(globalpath, "01_astro")), winslash = "/"))
```

Parameter transformations:

```{r}
parambounds <- list(
  Psi0 = c(1, 200),
  gamma = c(0.3, 0.7),
  alpha = c(3.0, 3.7),
  beta = c(-0.5, 1)
)

get_Psi0 <- function(theta) {
  # theta = -log( (Psi0 - 1) / (200 - 1) )
  (parambounds$Psi0[2] - parambounds$Psi0[1]) * exp(-exp(theta)) + parambounds$Psi0[1]
}

get_theta1 <- function(Psi0) {
  log(-log((Psi0 - parambounds$Psi0[1]) / (parambounds$Psi0[2] - parambounds$Psi0[1])))
}

get_gamma <- function(theta) {
  # theta = -log( (gamma - 0.3) / (0.7 - 0.3) )
  (parambounds$gamma[2] - parambounds$gamma[1]) * exp(-exp(theta)) + parambounds$gamma[1]
}

get_theta2 <- function(gamma) {
  log(-log((gamma - parambounds$gamma[1]) / (parambounds$gamma[2] - parambounds$gamma[1])))
}

get_alpha <- function(theta) {
  # theta = log(alpha - 3)
  exp(theta) + parambounds$alpha[1]
}

get_theta3 <- function(alpha) {
  log(alpha - parambounds$alpha[1])
}

get_beta <- function(theta) {
  # theta = -log( (beta - (-0.5)) / (1 - (-0.5)) )
  (parambounds$beta[2] - parambounds$beta[1]) * exp(-exp(theta)) + parambounds$beta[1]
}

get_theta4 <- function(beta) {
  log(-log((beta - parambounds$beta[1]) / (parambounds$beta[2] - parambounds$beta[1])))
}
```

# Optimisation

Optimisation using IPOPT replaced by optimisation using `nloptr`:

```{r}
nloptr_objective <- function(theta) ff$fn(theta)
nloptr_objective_gradient <- function(theta) ff$gr(theta)
```

We use box constraints, to improve stability of optimization:

```{r}
lowerbounds <- c(
  get_theta1(parambounds$Psi0[2] - 0.001),
  get_theta2(parambounds$gamma[2] - 0.001),
  get_theta3(parambounds$alpha[1] + 0.001),
  get_theta4(parambounds$beta[2] - 0.001)
)

upperbounds <- c(
  get_theta1(parambounds$Psi0[1] + 1),
  get_theta2(parambounds$gamma[1] + 0.01),
  get_theta3(parambounds$alpha[2] - 0.01),
  get_theta4(parambounds$beta[1] + 0.01)
)
```

Start optimisation at the midpoint:

```{r}
thetastart <- (upperbounds + lowerbounds) / 2
thetastart_list <- list(theta1 = thetastart[1], theta2 = thetastart[2], theta3 = thetastart[3], theta4 = thetastart[4])

data.frame(
  "param" = paste0("theta", 1:4),
  lowerbound = lowerbounds,
  upperbound = upperbounds,
  thetastart = thetastart
) %>%
  ggplot(aes(ymin = lowerbound, y = thetastart, ymax = upperbound, x = param)) +
    geom_pointrange() +
    coord_flip() +
    labs(x = "", y = "") +
    theme_minimal()
```

Nonlinear constraints, specified as a function:

```{r}
nloptr_nonlinear_constraints <- function(theta) -1 * Es$fn(theta)

nloptr_nonlinear_constraints_jacobian <- function(theta) {
  J <- -1 * Es$gr(theta)
  as.matrix(J)
}

nonlinear_lowerbound <- rep(0, nrow(gcdata) + 2)
nonlinear_upperbound <- rep(Inf, nrow(gcdata) + 2)
```

Objective function and its derivatives:

```{r}
ff <- MakeADFun(
  data = gcdatalist,
  parameters = thetastart_list,
  DLL = "01_astro",
  ADreport = FALSE,
  silent = TRUE
)
```

Nonlinear constraints and their Jacobian:

```{r}
Es <- MakeADFun(
  data = gcdatalist,
  parameters = thetastart_list,
  DLL = "01_astro",
  ADreport = TRUE,
  silent = TRUE
)
```

Check starting value satisfies constraints.
Note that constraints in `nloptr` are of the form $g(x) \leq 0$.

```{r}
stopifnot(all(nloptr_nonlinear_constraints(thetastart) <= 0))

tm <- Sys.time()

tmp <- capture.output(
  nloptr_result <- nloptr::nloptr(
    x0 = thetastart,
    eval_f = nloptr_objective,
    eval_grad_f = nloptr_objective_gradient,
    lb = lowerbounds,
    ub = upperbounds,
    eval_g_ineq = nloptr_nonlinear_constraints,
    eval_jac_g_ineq = nloptr_nonlinear_constraints_jacobian,
    opts = list("algorithm" = "NLOPT_LD_MMA", obj_scaling_factor = 1, "xtol_rel" = 1.0e-3)
  )
)

optruntime <- difftime(Sys.time(), tm, units = "secs")
cat("Run time for mass model optimisation:", optruntime, "seconds.\n")
```

Check that the optimisation results look right (based on Figure 2 of @bilodeau2021stochastic).

```{r}
get_Psi0(nloptr_result$solution[1])
get_gamma(nloptr_result$solution[2])
get_alpha(nloptr_result$solution[3])
get_beta(nloptr_result$solution[4])
```

# AGHQ

Create the template for AGHQ:

```{r message=FALSE}
ffa <- list(
  fn = function(x) -1 * ff$fn(x),
  gr = function(x) -1 * ff$gr(x),
  he = function(x) -1 * ff$he(x)
)

useropt <- list(
  ff = ffa,
  mode = nloptr_result$solution,
  hessian = ff$he(nloptr_result$solution)
)

#' Fit it with "reuse" first, then correct marginals later
cntrl <- aghq::default_control(negate = TRUE)

#' Do the quadrature
tm <- Sys.time()

k_dense <- 5
astroquad <- aghq::aghq(ff, k = k_dense, thetastart, optresults = useropt, control = cntrl)

#' Correct the marginals except for beta, which gave NA due to the constraints
for (j in 1:3) astroquad$marginals[[j]] <- marginal_posterior(astroquad, j, method = "correct")
quadruntime <- difftime(Sys.time(), tm, units = "secs")

cat("Run time for mass model quadrature:", quadruntime ,"seconds.\n")

#' Total runtime for AGHQ
optruntime + quadruntime
```

# EB

```{r message=FALSE}
tm <- Sys.time()

k_eb <- 1
astroebquad <- aghq::aghq(ff, k = k_eb, thetastart, optresults = useropt, control = cntrl)

#' Correct the marginals except for beta, which gave NA due to the constraints
for (j in 1:3) astroebquad$marginals[[j]] <- marginal_posterior(astroebquad, j, method = "correct")
quadebruntime <- difftime(Sys.time(), tm, units = "secs")

cat("Run time for mass model quadrature:", quadebruntime ,"seconds.\n")

#' Total runtime for EB
optruntime + quadebruntime
```

# PCA-AGHQ

```{r message=FALSE}
C <- Matrix::forceSymmetric(solve(useropt$hessian))
eigenC <- eigen(C)
lambda <- eigenC$values

data.frame(
  n = 1:length(lambda),
  tv = cumsum(lambda / sum(lambda))
) %>% ggplot(aes(x = n, y = tv)) +
  geom_point() +
  geom_hline(yintercept = 0.9, col = "grey", linetype = "dashed") +
  annotate("text", x = 3, y = 0.89, label = "90% of total variation explained", col = "grey") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = "PCA dimensions included", y = "Total variation explained") +
  theme_minimal()

k_pca <- 5
s <- 1 #' Chosen based on the Scree plot
m <- length(useropt$mode)
levels <- c(rep(k_pca, s), rep(1, m - s))
pca_grid <- mvQuad::createNIGrid(dim = m, type = "GHe", level = levels)
mvQuad::rescale(pca_grid, m = useropt$mode, C = C, dec.type = 1)

#' Check that the PCA grid has the correct number of nodes
stopifnot(nrow(mvQuad::getNodes(pca_grid)) == k_pca^s)

tm <- Sys.time()

astropcaquad <- local_aghq(ff, k_pca, thetastart, optresults = useropt, basegrid = pca_grid, adapt = FALSE, control = cntrl)

#' Correct the marginals except for beta, which gave NA due to the constraints
for (j in 1:3) astropcaquad$marginals[[j]] <- marginal_posterior(astropcaquad, j, method = "correct")
pcaquadruntime <- difftime(Sys.time(), tm, units = "secs")

cat("Run time for mass model quadrature:", pcaquadruntime ,"seconds.\n")

#' Total runtime for PCA-AGHQ
optruntime + pcaquadruntime
```

# NUTS

```{r}
stanmod <- tmbstan(
  ff,
  chains = 4,
  cores = 4,
  iter = 1e04,
  warmup = 1e03,
  init = nloptr_result$solution,
  seed = 48645,
  algorithm = "NUTS"
)

# Time
get_elapsed_time(stanmod)

standata <- as.data.frame(stanmod)
standata$Psi0 <- get_Psi0(standata[ ,1])
standata$gamma <- get_gamma(standata[ ,2])
standata$alpha <- get_alpha(standata[ ,3])
standata$beta <- get_beta(standata[ ,4])
```

# Comparison

## Plots

```{r warning=FALSE}
#' Manual plots
parambounds <- list(
  Psi0 = c(1, 200),
  gamma = c(0.3, 0.7),
  alpha = c(3.0, 3.7),
  beta = c(-0.5, 1)
)

#' Add small buffers for stability
get_theta2_robust <- function(gamma) {
  log(-log((gamma - parambounds$gamma[1] + 1e-03) / (parambounds$gamma[2] - parambounds$gamma[1] + 1e-03)))
}

get_theta3_robust <- function(alpha) log(alpha - parambounds$alpha[1] + 1e-03)

#' Transformed PDFs
translist1 <- make_transformation(totheta = get_theta1, fromtheta = get_Psi0)
translist2 <- make_transformation(totheta = get_theta2_robust, fromtheta = get_gamma)
translist3 <- make_transformation(totheta = get_theta3_robust, fromtheta = get_alpha)
translist4 <- make_transformation(totheta = get_theta4, fromtheta = get_beta)

Psi0pdf <- compute_pdf_and_cdf(astroquad$marginals[[1]], translist1)
gammapdf <- compute_pdf_and_cdf(astroquad$marginals[[2]], translist2)
alphapdf <- compute_pdf_and_cdf(astroquad$marginals[[3]], translist3, interpolation = "polynomial")
betapdf <- compute_pdf_and_cdf(astroquad$marginals[[4]], translist4)

Psi0ebpdf <- compute_pdf_and_cdf(astroebquad$marginals[[1]], translist1)
gammaebpdf <- compute_pdf_and_cdf(astroebquad$marginals[[2]], translist2)
alphaebpdf <- compute_pdf_and_cdf(astroebquad$marginals[[3]], translist3, interpolation = "polynomial")
betaebpdf <- compute_pdf_and_cdf(astroebquad$marginals[[4]], translist4)

Psi0pcapdf <- compute_pdf_and_cdf(astropcaquad$marginals[[1]], translist1)
gammapcapdf <- compute_pdf_and_cdf(astropcaquad$marginals[[2]], translist2)
alphapcapdf <- compute_pdf_and_cdf(astropcaquad$marginals[[3]], translist3, interpolation = "polynomial")
betapcapdf <- compute_pdf_and_cdf(astropcaquad$marginals[[4]], translist4)

Psi0prior <- function(Psi0) dunif(Psi0, parambounds$Psi0[1], parambounds$Psi0[2], log = FALSE)
gammaprior <- function(gamma) dunif(gamma, parambounds$gamma[1], parambounds$gamma[2], log = FALSE)
alphaprior <- function(alpha) dgamma(alpha - parambounds$alpha[1], shape = 1, rate = 4.6, log = FALSE)
betaprior <- function(beta) dunif(beta, parambounds$beta[1], parambounds$beta[2], log = FALSE)

(Psi0_postplot <- Psi0pdf %>%
  ggplot(aes(x = transparam, y = pdf_transparam)) +
  geom_histogram(
    data = standata,
    mapping = aes(x = Psi0, y = after_stat(density)),
    bins = 50,
    colour = "white",
    fill = "grey"
  ) +
  geom_line() +
  geom_line(
    data = Psi0pcapdf,
    inherit.aes = FALSE,
    mapping = aes(x = transparam, y = pdf_transparam),
    colour = "blue",
  ) +
  stat_function(fun = Psi0prior, linetype = "dashed") +
  labs(title = "", x = "", y = "Density") +
  scale_x_continuous(breaks = seq(10, 60, by = 5)) +
  coord_cartesian(xlim = c(24, 43)) +
  theme_minimal())

(gamma_postplot <- gammapdf %>%
  ggplot(aes(x = transparam, y = pdf_transparam)) +
  geom_histogram(
    data = standata,
    mapping = aes(x = gamma, y = after_stat(density)),
    bins = 100,
    colour = "white",
    fill = "grey"
  ) +
  geom_line() +
  geom_line(
    data = gammapcapdf,
    mapping = aes(x = transparam, y = pdf_transparam),
    colour = "blue",
  ) +
  stat_function(fun = gammaprior, linetype = "dashed") +
  labs(title = "", x = "", y = "Density") +
  scale_x_continuous(breaks = seq(0, 0.44, by = 0.02)) +
  coord_cartesian(xlim = c(0.3, 0.4)) +
  theme_minimal())

(alpha_postplot <- alphapdf %>%
  ggplot(aes(x = transparam, y = pdf_transparam)) +
  geom_histogram(
    data = standata,
    mapping = aes(x = alpha, y = after_stat(density)),
    bins = 2000,
    colour = "white",
    fill = "grey"
  ) +
  geom_line() +
  geom_line(
    data = alphapcapdf,
    mapping = aes(x = transparam, y = pdf_transparam),
    colour = "blue",
  ) +
  stat_function(fun = alphaprior, linetype = "dashed") +
  labs(title = "",x = "",y = "Density") +
  scale_x_continuous(breaks = seq(3, 4, by = 0.02)) +
  coord_cartesian(xlim = c(3, 3.05)) +
  theme_minimal())

(beta_postplot <- betapdf %>%
  ggplot(aes(x = transparam, y = pdf_transparam)) +
  geom_histogram(
    data = standata,
    mapping = aes(x = beta, y = after_stat(density)),
    bins = 50,
    colour = "white",
    fill = "grey"
  ) +
  geom_line() +
  geom_line(
    data = betapcapdf,
    mapping = aes(x = transparam, y = pdf_transparam),
    colour = "blue",
  ) +
  stat_function(fun = betaprior, linetype = "dashed") +
  labs(title = "",x = "",y = "Density") +
  scale_x_continuous(breaks = seq(-0.5, 1, by = 0.1)) +
  coord_cartesian(xlim = c(-0.3, 0.4)) +
  theme_minimal())
```

## Normalising constant

```{r}
data.frame(
  method = c("AGHQ", "PCA-AGHQ", "EB"),
  lognormconst = c(astroquad$normalized_posterior$lognormconst, astropcaquad$normalized_posterior$lognormconst, astroebquad$normalized_posterior$lognormconst)
) %>%
  gt::gt()
```

## KS tests

```{r warning=FALSE}
kstable <- function(quad) {
  aghqpostsamps_spline <- sample_marginal(quad, nrow(standata), interpolation = "spline")
  aghqpostsamps_poly <- sample_marginal(quad, nrow(standata), interpolation = "polynomial")

  suppressWarnings({
    kstable <- data.frame(
      Psi0 = ks.test(standata[[1]], aghqpostsamps_poly[[1]])$statistic,
      gamma = ks.test(standata[[2]], aghqpostsamps_poly[[2]])$statistic,
      alpha = ks.test(standata[[3]], aghqpostsamps_poly[[3]])$statistic,
      beta = ks.test(standata[[4]], aghqpostsamps_poly[[4]])$statistic
    )
  })

  return(kstable)
}

kstable_astroquad <- kstable(astroquad) %>%
  pivot_longer(cols = everything()) %>%
  mutate(method = "AGHQ", k = k_dense, s = NA, total_points = k^4)

#' Not working (yet?)
# kstable_astropcaquad <- kstable(astropcaquad) %>%
#   pivot_longer(cols = everything()) %>%
#   mutate(method = "AGHQ-PCA", k = k_pca, s = s, total_points = k^s)

bind_rows(kstable_astroquad) %>%
  gt::gt()

bind_rows(kstable_astroquad) %>%
  ggplot(aes(x = name, y = value, col = method)) +
    geom_jitter(height = 0, width = 0.05) +
    scale_y_continuous(limits = c(0, 0.05)) +
    scale_colour_manual(values = multi.utils::cbpalette()) +
    labs(x = "Parameter", y = "KS(method, NUTS)", col = "Method") +
    theme_minimal()
```

# Original computing environment {-}

```{r}
sessionInfo()
```

# Bibliography {-}