### 2024-12-07 src/docs_paper-jtb/paper.Rmd
Add JTB template

---
title: Fast approximate Bayesian inference of HIV indicators using PCA adaptive Gauss-Hermite quadrature
author:
  - name: Adam Howes
    email: ath19@ic.ac.uk
    affiliation: A
    correspondingauthor: true
  - name: Alex Stringer
    email: alex.stringer@uwaterloo.ca
    affiliation: B
  - name: Seth R. Flaxman
    email: seth.flaxman@cs.ox.ac.uk
    affiliation: C
  - name: Jeffrey W. Imai-Eaton
    email: jeaton@hsph.harvard.edu
    affiliation: D
address:
  - code: A
    organization: Department of Mathematics
    country: Imperial College London
  - code: B
    organization: Department of Statistics and Actuarial Science
    country: University of Waterloo
  - code: C
    organization: Department of Computer Science
    country: University of Oxford
  - code: D
    organization: Harvard T.H. Chan School of Public Health
    country: Harvard University
abstract: |
  This is the abstract.
keywords: 
  - Bayesian statistics
  - spatial statistics
  - evidence synthesis
  - small-area estimation
  - approximate inference
  - INLA
  - AGHQ
  - HIV epidemiology
journal: "Journal of Theoretical Biology"
date: "`r Sys.Date()`"
linenumbers: false
numbersections: true
bibliography: citations.bib
biblio-style: elsarticle-harv # author year style for natbib - use 'elsarticle-num' or 'elsarticle-num-names' for numbered scheme
classoption: preprint, 3p, authoryear # remove authoryear is not using `elsarticle-harv`
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---

# References {-}