### 2024-12-29 src/docs_paper-jtb/preamble.tex

\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
\RequirePackage{bm}
% \RequirePackage{enumerate}
% \RequirePackage{enumitem}
\RequirePackage{tabularx}
% \RequirePackage{adjustbox}
% \RequirePackage{tikz}
% \RequirePackage{bbm}
% \RequirePackage{lscape}
% \RequirePackage{booktabs}
% \RequirePackage{longtable}
% \RequirePackage{lscape}
% \RequirePackage{pifont}
% \usepackage{footnote}
% \usepackage{algorithm}
% \usepackage{algpseudocode}

\newcommand{\Sc}{\mathcal{S}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\m}{\mathbf{m}}
\newcommand{\bk}{\mathbf{k}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\z}{\mathbf{z}}
\newcommand{\bb}{\mathbf{b}}
\newcommand{\Hb}{\mathbf{H}}
\newcommand{\bphi}{\bm{\phi}}
\newcommand{\bpsi}{\bm{\psi}}
\newcommand{\brho}{\bm{\rho}}
\newcommand{\btheta}{\bm{\theta}}
\newcommand{\bmeta}{\bm{\eta}}
\newcommand{\bmu}{\bm{\mu}}

\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\usepackage{amsmath}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}