### 2025-05-10 src/docs_paper-jtb/revisions.Rmd
Start with revisions (a typo!)

---
title: Response to reviewers
author:
  - Adam Howes^[Department of Mathematics, Imperial College London]
  - Alex Stringer^[Department of Statistics and Actuarial Science, University of Waterloo]
  - Seth R. Flaxman^[Department of Computer Science, Oxford University]
  - Jeffrey W. Imai-Eaton^[Harvard T.H. Chan School of Public Health, Harvard University]
output:
  bookdown::pdf_document2:
    toc: no
    number_sections: yes
    keep_tex: yes
bibliography: citations.bib
header-includes:
  \usepackage{xr} \externaldocument{paper}
---

```{r echo = FALSE, message = FALSE}
library(dplyr)

options(scipen = 100)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dpi = 320,
  cache = TRUE,
  out.width = "95%",
  fig.align = "center",
  fig.pos = "!h",
  out.extra = ""
)

knitr::opts_chunk$set()
```

Thank you for taking the time to read and review our paper "Fast approximate Bayesian inference of HIV indicators using PCA adaptive Gauss-Hermite quadrature".
In what follows we address your comments and suggestions point by point.
We appreciate the feedback to improve the work and have made the relevant changes to the manuscript.

\clearpage

# Reviewer 1 {-}

> Typos spotted

Thank you for spotting these two typographic errors.
We have corrected them in the manuscript.

# Reviewer 2 {-}

# References {#references .unnumbered}