### 2025-05-26 src/naomi-simple_point-estimates/point-estimates.Rmd

---
title: "Point estimates comparison for the simplified Naomi model"
author:
- name: Adam Howes
output:
  bookdown::html_document2:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Background**  We have run the simplified Naomi model using a range of inference methods: `TMB`, `aghq`, and `tmbstan`.

  **Task** In this report, we compare the accuracy of the posterior distributions obtained from these inference methods using point estimates.
---

# Background

We compare the inference results from `TMB`, `aghq`, and `tmbstan`.
Import these inference results as follows:

```{r}
tmb <- readRDS("depends/tmb.rds")
aghq <- readRDS("depends/aghq.rds")
tmbstan <- readRDS("depends/tmbstan.rds")
```

# Data manipulation

We're interested in the latent field and certain model outputs (prevalence, incidence and treatment coverage):

```{r}
latent_pars <- unique(names(tmb$fit$par.full))[!(unique(names(tmb$fit$par.full)) %in% names(tmb$fit$par))]
output_pars <- c("rho_t1_out", "lambda_t1_out", "alpha_t1_out")

type_key <- data.frame(
  par = c(latent_pars, output_pars),
  type = c(rep("latent", times = length(latent_pars)), rep("output", times = length(output_pars)))
)
```

Extract the mean and standard deviation for each parameter for each inference method, then bind them together:

```{r}
df_tmb <- lapply(tmb$fit$sample[type_key$par], function(row) {
  data.frame("mean" = rowMeans(row), "sd" = matrixStats::rowSds(row)) %>%
    tibble::rowid_to_column("id")
  }) %>%
  dplyr::bind_rows(.id = "par") %>%
  mutate(method = "GEB")

df_aghq <- lapply(aghq$quad$sample[type_key$par], function(row) {
  data.frame("mean" = rowMeans(row), "sd" = matrixStats::rowSds(row)) %>%
    tibble::rowid_to_column("id")
  }) %>%
  dplyr::bind_rows(.id = "par") %>%
  mutate(method = "GPCA-AGHQ")

df_tmbstan <- lapply(tmbstan$mcmc$sample[type_key$par], function(row) {
  data.frame("mean" = rowMeans(row), "sd" = matrixStats::rowSds(row)) %>%
    tibble::rowid_to_column("id")
  }) %>%
  dplyr::bind_rows(.id = "par") %>%
  mutate(method = "NUTS")

df <- bind_rows(df_tmb, df_aghq, df_tmbstan) %>%
  left_join(
    type_key,
    by = "par"
  )

head(df)
```

Pivot so that the data is ready to plot:

```{r}
df_plot <- df %>%
  pivot_longer(cols = c("sd", "mean"), names_to = "indicator", values_to = "estimate") %>%
  pivot_wider(values_from = "estimate", names_from = "method") %>%
  pivot_longer(cols = c("GEB", "GPCA-AGHQ"), names_to = "method", values_to = "approximate") %>%
  rename("truth" = "NUTS") %>%
  mutate(
    indicator = fct_recode(indicator, "Posterior mean" = "mean", "Posterior SD" = "sd"),
    method = fct_relevel(method, "GEB", "GPCA-AGHQ")
  )

head(df_plot)
```
Calculate the root mean square and mean absolute errors between the approximate methods and NUTS (taken to be the truth):

```{r}
write_csv(df_plot, "point-estimates.csv")

df_metrics <- df_plot %>%
  group_by(method, indicator, type) %>%
  summarise(
    rmse = sqrt(mean((truth - approximate)^2)),
    mae = mean(abs(truth - approximate))
  ) %>%
  ungroup()

head(df_metrics)
```

Add difference in metric between GEB and GPCA-AGHQ:

```{r}
df_metrics_pct <- df_metrics %>%
  split(~ indicator + type) %>%
  lapply(function(x) {
    x %>% summarise(
      rmse_diff = 100 * diff(rmse) / max(rmse),
      mae_diff = 100 * diff(mae) / max(mae)
    )
  }) %>%
  bind_rows(.id = "id") %>%
  separate(id, c("indicator", "type"), sep = "\\.")

#' Add as a naming column for plot
df_metrics <- df_metrics %>%
  left_join(
    df_metrics_pct,
    by = c("indicator", "type")
  ) %>%
  mutate(
    label = ifelse(
      method == "GPCA-AGHQ",
      paste0("RMSE: ", signif(rmse, 2), " (", signif(rmse_diff, 2), "%)", "\nMAE: ", signif(mae, 2), " (", signif(mae_diff, 2), "%)"),
      paste0("RMSE: ", signif(rmse, 2), "\nMAE: ", signif(mae, 2))
    )
  )

write_csv(df_metrics, "mean-sd.csv")
```

# Plots

## Latent field

```{r}
df_plot_latent <- filter(df_plot, type == "latent")
df_metrics_latent <- filter(df_metrics, type == "latent")

df_plot_latent %>%
  ggplot(aes(x = truth, y = approximate - truth)) +
  geom_point(shape = 1, alpha = 0.4) +
  facet_grid(indicator ~ method) +
  geom_abline(intercept = 0, slope = 0, col = "#E69F00", linetype = "dashed") +
  geom_text(data = df_metrics_latent, aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5) +
  labs(x = "Estimate", y = "Absolute difference (NUTS - method)") +
  theme_minimal()

ggsave("mean-sd-latent.png", h = 6, w = 6.25)
```

Version split into two for presentations

```{r}
jitter_amount <- 0.02

mean_plot_latent <- df_plot_latent %>%
  filter(indicator == "Posterior mean") %>%
  ggplot(aes(x = truth, y = approximate - truth)) +
  geom_jitter(shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  lims(y = c(-0.4, 0.4)) +
  facet_grid(indicator ~ method) +
  geom_abline(intercept = 0, slope = 0, col = "#E69F00", linetype = "dashed") +
  geom_text(
    data = filter(df_metrics_latent, indicator == "Posterior mean"),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "Estimate", y = "Absolute difference (NUTS - method)") +
  theme_minimal()

ggsave("mean-latent.png", mean_plot_latent, h = 4, w = 6.25)

sd_plot_latent <- df_plot_latent %>%
  filter(indicator == "Posterior SD") %>%
  ggplot(aes(x = truth, y = approximate - truth)) +
  geom_jitter(shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  lims(y = c(-0.6, 0.6)) +
  facet_grid(indicator ~ method) +
  geom_abline(intercept = 0, slope = 0, col = "#E69F00", linetype = "dashed") +
  geom_text(
    data = filter(df_metrics_latent, indicator == "Posterior SD"),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "Estimate", y = "Absolute difference (NUTS - method)") +
  theme_minimal()

ggsave("sd-latent.png", sd_plot_latent, h = 4, w = 6.25)

combine_two_plots <- function(plot1, plot2) {
  plot1alt <- plot1 +
    theme(legend.position = "none")

  plot2alt <- plot2 +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    legend.position = "none"
  )

  y_axis <- ggplot(data.frame(l = plot1$labels$y, x = 1, y = 1)) +
    geom_text(aes(x, y, label = l), angle = 90) +
    theme_void() +
    coord_cartesian(clip = "off")

  plot1alt$labels$x <- ""
  plot1alt$labels$y <- plot2alt$labels$y <- ""

  y_axis + (plot1alt / plot2alt) +
    plot_layout(widths = c(1, 35))
}

combine_two_plots(mean_plot_latent, sd_plot_latent)

ggsave("mean-sd-alt-latent.png", h = 6, w = 6.25)
```

Where are the greatest errors occurring for the mean?

```{r}
df_plot_latent %>%
  filter(indicator == "Posterior mean") %>%
  mutate(diff = abs(approximate - truth)) %>%
  arrange(desc(diff)) %>%
  select(indicator, par, id, method, truth, approximate, diff) %>%
  DT::datatable()
```

What about for the standard deviations?

```{r}
df_plot_latent %>%
  filter(indicator == "Posterior SD") %>%
  mutate(diff = abs(approximate - truth)) %>%
  arrange(desc(diff)) %>%
  select(indicator, par, id, method, truth, approximate, diff) %>%
  DT::datatable()
```

## Outputs

```{r}
df_plot_output <- filter(df_plot, type == "output")
df_metrics_output <- filter(df_metrics, type == "output")

df_plot_output <- df_plot_output %>%
  mutate(par = fct_recode(par, "HIV prevalence" = "rho_t1_out", "ART coverage" = "alpha_t1_out", "HIV incidence" = "lambda_t1_out"))

df_plot_output %>%
  ggplot() +
  geom_point(aes(x = truth, y = approximate - truth, color = par), shape = 1, alpha = 0.4) +
  scale_color_manual(values = c("#56B4E9","#009E73", "#CC79A7")) +
  facet_grid(indicator ~ method) +
  geom_abline(intercept = 0, slope = 0, col = "#E69F00", linetype = "dashed") +
  geom_text(data = df_metrics_output, aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5) +
  labs(x = "Estimate", y = "Absolute difference (NUTS - method)", col = "Output") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("mean-sd-output.png", h = 6, w = 6.25)
```

Version split into two for presentations

```{r}
jitter_amount <- 0

mean_plot_output <- df_plot_output %>%
  filter(indicator == "Posterior mean") %>%
  ggplot() +
  geom_jitter(aes(x = truth, y = approximate - truth, color = par), shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  scale_color_manual(values = c("#56B4E9","#009E73", "#CC79A7")) +
  facet_grid(indicator ~ method) +
  geom_abline(intercept = 0, slope = 0, col = "#E69F00", linetype = "dashed") +
  geom_text(
    data = filter(df_metrics_output, indicator == "Posterior mean"),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "Estimate", y = "Absolute difference (NUTS - method)", col = "Output") +
  theme_minimal()

ggsave("mean-output.png", mean_plot_output, h = 4, w = 6.25)

sd_plot_output <- df_plot_output %>%
  filter(indicator == "Posterior SD") %>%
  ggplot() +
  geom_jitter(aes(x = truth, y = approximate - truth, color = par), shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  scale_color_manual(values = c("#56B4E9","#009E73", "#CC79A7")) +
  facet_grid(indicator ~ method) +
  geom_abline(intercept = 0, slope = 0, col = "#E69F00", linetype = "dashed") +
  geom_text(
    data = filter(df_metrics_output, indicator == "Posterior SD"),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "Estimate", y = "Absolute difference (NUTS - method)", col = "Output") +
  theme_minimal()

ggsave("sd-output.png", sd_plot_output, h = 4, w = 6.25)

combine_two_plots(mean_plot_output, sd_plot_output)

ggsave("mean-sd-alt-output.png", h = 6, w = 6.25)
```

Version split by output type:

```{r}
df_metrics_output_split <- df_plot_output %>%
  group_by(method, indicator, type, par) %>%
  summarise(
    rmse = sqrt(mean((truth - approximate)^2)),
    mae = mean(abs(truth - approximate))
  ) %>%
  ungroup()

df_metrics_output_split_pct <- df_metrics_output_split %>%
  split(~ indicator + type + par) %>%
  lapply(function(x) {
    x %>% summarise(
      rmse_diff = 100 * diff(rmse) / max(rmse),
      mae_diff = 100 * diff(mae) / max(mae)
    )
  }) %>%
  bind_rows(.id = "id") %>%
  separate(id, c("indicator", "type", "par"), sep = "\\.")

df_metrics_output_split <- df_metrics_output_split %>%
  left_join(
    df_metrics_output_split_pct,
    by = c("indicator", "type", "par")
  ) %>%
  mutate(
    label = ifelse(
      method == "GPCA-AGHQ",
      paste0("RMSE: ", signif(rmse, 2), " (", signif(rmse_diff, 2), "%)", "\nMAE: ", signif(mae, 2), " (", signif(mae_diff, 2), "%)"),
      paste0("RMSE: ", signif(rmse, 2), "\nMAE: ", signif(mae, 2))
    )
  )

plot_outputs <- function(.indicator, .par, col) {
  df_plot_output %>%
  filter(indicator == .indicator, par == .par) %>%
  ggplot() +
  geom_jitter(aes(x = truth, y = approximate - truth, color = par), shape = 1, alpha = 0.4, width = jitter_amount, height = jitter_amount) +
  scale_color_manual(values = c(col, "white", "white")) +
  facet_grid(indicator + par ~ method, labeller = labeller(indicator = function(x) {rep("", length(x))})) +
  geom_abline(intercept = 0, slope = 0, col = "#E69F00", linetype = "dashed") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 4)) +
  geom_text(
    data = filter(df_metrics_output_split, indicator == .indicator, par == .par),
    aes(x = -Inf, y = Inf, label = label), size = 3, hjust = 0, vjust = 1.5
  ) +
  labs(x = "Estimate", y = "Absolute difference (NUTS - method)", col = "Output") +
  theme_minimal() +
  theme(legend.position = "none", panel.spacing.x = unit(6, "mm"))
}

mean_prev_plot <- plot_outputs(.indicator = "Posterior mean", .par = "HIV prevalence", col = "#CC79A7")
mean_inc_plot <- plot_outputs(.indicator = "Posterior mean", .par = "HIV incidence", col = "#009E73")
mean_art_plot <- plot_outputs(.indicator = "Posterior mean", .par = "ART coverage", col = "#56B4E9")
sd_prev_plot <- plot_outputs(.indicator = "Posterior SD", .par = "HIV prevalence", col = "#CC79A7")
sd_inc_plot <- plot_outputs(.indicator = "Posterior SD", .par = "HIV incidence", col = "#009E73")
sd_art_plot <- plot_outputs(.indicator = "Posterior SD", .par = "ART coverage", col = "#56B4E9")

combine_three_plots <- function(plot1, plot2, plot3, subtitle) {

  plot1alt <- plot1 + theme(legend.position = "none") + labs(subtitle = subtitle)

  plot2alt <- plot2 +
    theme(
      strip.background = element_blank(),
      strip.text.x = element_blank(),

      legend.position = "none"
    )

  plot3alt <- plot3 +
    theme(
      strip.background = element_blank(),
      strip.text.x = element_blank(),
      legend.position = "none"
    )

  y_axis <- ggplot(data.frame(l = plot1$labels$y, x = 1, y = 1)) +
    geom_text(aes(x, y, label = l), angle = 90) +
    theme_void() +
    coord_cartesian(clip = "off")

  plot1alt$labels$x <- plot2alt$labels$x <- ""
  plot1alt$labels$y <- plot2alt$labels$y <- plot3alt$labels$y <- ""

  three_plots <- cowplot::plot_grid(plot1alt, plot2alt, plot3alt, ncol = 1, rel_heights = c(1.25, 1, 1))
  cowplot::plot_grid(y_axis, three_plots, ncol = 2, rel_widths = c(1, 35))
}

combine_three_plots(mean_prev_plot, mean_inc_plot, mean_art_plot, subtitle = "Posterior mean")
ggsave("mean-alt-output.png", h = 7, w = 6.25)

combine_three_plots(sd_prev_plot, sd_inc_plot, sd_art_plot, subtitle = "Posterior SD")
ggsave("sd-alt-output.png", h = 7, w = 6.25)
```

# Original computing environment {-}

```{r}
sessionInfo()
```

# Bibliography {-}