### 2025-08-20 src/docs_paper-jtb/revisions.Rmd
JIE suggestions

---
title: 
author:
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: true
    keep_tex: true
    includes:
      in_header: preamble.tex
    extra_dependencies: awesomebox
bibliography: citations.bib
header-includes:
  \usepackage{xr} \externaldocument{paper}
---

```{r echo = FALSE, message = FALSE}
library(dplyr)

options(scipen = 100)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  dpi = 320,
  cache = TRUE,
  out.width = "95%",
  fig.align = "center",
  fig.pos = "!h",
  out.extra = ""
)

knitr::opts_chunk$set()
```

\clearpage

Dear Dr. XXXX,

Many thanks for your consideration of our paper "Fast approximate Bayesian inference of HIV indicators using PCA adaptive Gauss-Hermite quadrature" and to the reviewers for their thoughtful comments on the manuscript.

Below we address the reviewers' comments and suggestions point by point. While updating and extending the simulations to address some of the reviewers' comments, we noticed that the results reported arose from two different versions of the model, with some from an alternative experimental algorithm that was not retained for the final method. Along with updating all results with longer MCMC chain simulations, we have now corrected the results in the manuscript to all be from the PCA grid described. This change impacts numerical results reported for the KS test, MMD, and exceedance probabilities, but does not change the overall conclusions or recommendations from the paper.

We appreciate the feedback to improve the work and have made the relevant changes to the manuscript as described below.

Sincerely,  
Adam Howes on behalf of Alex Stringer, Seth Flaxman, and Jeff Imai-Eaton

# Reviewer 1

\begin{reviewer}
\begin{itemize}
\item [R1.1] The class of models to which the present model is assigned, namely Extended Latent Gaussian Models (ELGMs), is very broad, and in my (perhaps rather curmudgeonly) opinion too broad to be of much value. On page 3 the authors write about the new mapping structure for ELGMs which is said to "relax the restriction that there is a one-to-one mapping between the mean response mu and structured additive predictor eta." and "Importantly, this mapping allows for the presence of non-linearity in the model." What exactly is meant by this? The original link function of LGMs can be non-linear of course, though it must sets up invertibility which is easy if it is one-to-one in that sense. Here the many-to-one refers to a mapping from multiple latent parameters to a combined latent parameter, so for example if we're talking about direct summation then yes we can have equality on a set of zero-measure for the many (the pull back of the 'one'), but I bet that for the model to work (in the sense of obeying the conditions of the ELGM convergence proofs) we couldn't, say, define the mapping such that its pull-back is sometimes a set of measure zero and sometimes a set of non-zero measure. It probably wouldn't work if we did a non-linear mapping like the product of the many. But who knows, because all this is left to the user of the ELGMs to figure out: that is, the original ELGM paper is constructed backwards: rather than defining a very strict model class with conditions that ensure a shared computability and set of theoretical results (like the LGM or GLM class) it is instead defined very loosely, meaning that finding out if it's computable using the given algorithms is left up to the user. Rant over. What can the authors do in the present paper? First, I would suggest to be clear on page 3 about what kind of mappings are being done in this model and see if these also match to a named smaller class within the ELGM class, and/or identify fuzzy matches to existing model structures of this form (such as Katie Wilson's "pointless spatial models"); especially those use in spatial statistics and/or epi.
\end{itemize}
\end{reviewer}

\textbf{Response:} We thank the reviewer for the thoughtful and insightful critique of the ELGM framing, and also for the concrete and practical suggestions for how we can address this to improve the current paper. We have added the following text in Section 2.2 with the aim of making the mapping between the latent field and structured additive predictor of the Naomi model clearer:

> Section 3 (e.g. Equations(8) to (11)) shows that the Naomi model involves many-to-one mappings of latent field components to the structured additive predictor for each response. Because of this, the Naomi model cannot be written as an LGM of the form described in Section 2.1, but can be written as an ELGM.

We are not aware of any named smaller class of ELGMs to which the Naomi model belongs.
The available convergence theory for ELGMs is very restrictive and it is not clear whether or how the available results apply to the Naomi model; accordingly, we do not make any claims about theoretical performance in the present paper.
Our empirical analysis shows that we obtain results that are practically similar to gold-standard NUTS results.

\begin{reviewer}
\begin{itemize}
\item [R1.2] I'm not an expert on the various flavours of NUTS algorithms but from my experience I have always been disappointed in the performance of tmbstan. Are there any comparisons between tmbstan and ordinary stan not conducted by the tmbstan package authors? Do people use tmbstan much in research? The software paper is well-cited but a lot of citations seem to be from other people just talking about software or HMC generally. Perhaps the best way to ally my concern that tmbstan might not be a good benchmark would be to check and explain in the paper whether its flavour of NUTS implementation is comparable to that now used in Stan? If not, perhaps see if a caveat is needed, something along the lines of 'tmbstan was tested because it's compatible with the idea of repurposing existing software (i.e., the original TMB template) but this doesn't mean that rebuilding the model from scratch in Stan, for instance, would create a sampler just as impractically slow'.
\end{itemize}
\end{reviewer}

\textbf{Response:} `tmbstan` passes the objective function and its gradient (calculated in `TMB`) directly to `rstan` via the [`op_matrix_vari`](https://mc-stan.org/math/classstan_1_1math_1_1op__matrix__vari.html).
As such, the NUTS algorithm used is exactly the same (as the same underlying `rstan` code is called.
It could be the case that there are speed differences between function evaluations in `TMB` and Stan.
We have added a sentence to the discussion noting this possible comparison:

> Additionally, direct implementation of Naomi within the Stan PPL could alter the speed by as much as plus or minus 50% based on experiements conducted by others [@monnahan2018no] and would give access to other algorithms implemented in Stan such as variational inference [@kucukelbir2015automatic] and Pathfinder [@zhang2022pathfinder].

\begin{reviewer}
\begin{itemize}
\item [R1.3] The equation for the fraction recently infected on page 17 puzzles me; though I don't come from the HIV field. It seems to me that this is a convenience 'guesstimation' used only as an output for end users, rather than being fitted to data or offering some kind structural constraint on the fits? Can these model outputs be pushed into a compartmental model later to give 'proper' incidence estimates, i.e., ones inferred from changes to prevalence and death rates? Also, presumably the epidemiology prescribes some tight bounds on the ANC offsets given the demography of birth and sexual partnership; are these pre-built into the priors for the model?
\end{itemize}
\end{reviewer}

\textbf{Response:} Some HIV surveys, including the example data from Malawi used in the analysis reported here, provide bio-marker data on whether or not HIV infection was recently acquired (approximately within the previous 4 months; the $\Omega_T$ parameter in Equation 7). This information on the proportion recently infected along with characteristics of the assay used to measure recent infection provides a direct survey estimator for HIV incidence, as derived by Kassanjee \textit{et al.} (https://pubmed.ncbi.nlm.nih.gov/22627902/). The formula in Equation 7 is a reformulation of the Kassanjee \textit{et al.} estimator to express a likelihood for the survey observations of the number recenlty infected, enabling Bayesian inference on the $\lambda_i$ parameters (HIV incidence rate). Details of the derivation of Equation 7 are described in the supplementary materials of Eaton \text{et al.} (and also Bao \text{et al.} (https://pubmed.ncbi.nlm.nih.gov/25406754/)). We have revised the manuscript here to clarify that Equation 7 is derived from the classical Kassanjee \textit{et al.} estimator:

> The proportion recently infected $\kappa_i$ is linked to HIV incidence via a reformulation of the Kassanjee \textit{et al.} biomarker-based incidence estimator...

The proposed approach by the reviewer to extend the model described here to a compartmental model that makes inference on HIV incidence from fitting to time series of HIV prevalence is indeed promising and an active area of research for our group and others. However, we feel this is separate from and outside the scope of the methodological inference questions addressed in this paper and describing this potential extension adds complexity that distracts from the primary research questions addressed here.

The reviewer is also correct to intuit that there are well characterised relationships between fertility and HIV infection that shape the pattern of HIV prevalence among pregnant women versus all women over the life course. However, at the ecological level the relationship can be somewhat more complex because of other confounding factors that affect both fertility and HIV risk, resulting in variation in the ANC versus population offsets across countries (for example, higher fertility and lower HIV prevalence in rural areas compared to urban areas). For this reason, we opted for a diffuse prior on the offset parameters, which in practice is well identified with the data available in applications.

\begin{reviewer}
\begin{itemize}
\item [R1.4] You anticipated my very clever idea for improving the performance of the PCA algorithm: namely, to consider the relevance of each dimension to a measure of uncertainty relevant to the end predictive purposes. Are calculations of this nature possible with TMB as it stands, or would one need to dig into the bowels of the c++ code and create a TMB package extension? I think this is an interesting area for future work.
\end{itemize}
\end{reviewer}

\textbf{Response:} We agree that this is an interesting area for future work. Regarding the question about whether this is possible within the existing `TMB` infrastructure, we believe it would be. We have added the following sentence to Section 6.1.2. discussing how:

> This may be achievable by applying the `ADREPORT` function to model outputs within the `TMB` template to obtain gradients.

\begin{reviewer}
\begin{itemize}
\item [R1.5] Finally, on page 17 it feels like Bayesian 'optimisation' is another natural competitor in this space, especially since it's quite straightforward to just run this kind of algorithm to explore hyper parameters until you get bored or run out of time, i.e., you get what you get from stopping early, unlike HMC for instance.
\end{itemize}
\end{reviewer}

\textbf{Response:} Our understanding of the goal of Bayesian optimisation (BO) is to find the maximum of a (black box) function, rather than to perform Bayesian inference.
(For example, one could imagine using BO to optimise the hyperparameters.)

\begin{reviewer}
\begin{itemize}
\item [R1.6] Typos spotted:
p2 "closer [to]"
p4. "hyperparmters"
\end{itemize}
\end{reviewer}

\textbf{Response:} We thank the reviewer for spotting these two typographic errors. We have corrected these in the manuscript.

# Reviewer 2

\begin{reviewer}
\begin{itemize}
\item [R2.1] The NUTS sampler was run for 4 chains for 100,000 iterations, with the first half discarded. This gives a lowest ESS of 208 and a largest Rhat of 1.021. This run took 79 hours. While this run is pretty reasonable for applied work, I do worry about it being used as a gold standard to compare against. Given that the authors are comparing the full posterior distribution, the tails are quite important which will be poorly characterised with ESS of 208. The Rhat of 1.021 is just about ok, so only slightly longer runs, but many more chains would, at fairly little expense, really improve the ESS. Can the model be run with 16 chains on an HPC or something?
\end{itemize}
\end{reviewer}

\textbf{Response:} We agree that running more iterations of the NUTS sampler is a good idea, and could improve the minimum ESS and accuracy of any comparisons, especially distributional ones.
We have rerun the NUTS sampler using a HPC with 18 chains (across 18 cores) with 40,000 per chain and a thinning factor of 20.
To improve the sampling also increased the `adapt_delta` parameter from the default of 0.8 to 0.95, and the `max_treedepth` parameter from the default of 10 to 12.
The resulting diagnostics are much improved. Updated sampler diagnostic results are reported in Section 5.1:

```{r}
naomi_nuts <- readRDS("depends/mcmc-out.rds")
```

> The Naomi model was challenging to efficiently sample from using NUTS via `tmbstan`.
We used 18 chains of 40000 iterations run on a high-performance computing cluster (Table \@ref(tab:inference-methods)).
The lowest resulting effective sample size (ESS) was `r round(naomi_nuts$ess_min, 0)` (2.5% quantile `r round(naomi_nuts$ess_lower, 0)`, 50% quantile `r round(naomi_nuts$ess_median, 0)`, and 97.5% quantile `r round(naomi_nuts$ess_upper, 0)`; Supplementary Figure 1A).
The largest resulting potential scale reduction factor ($\hat R$) was `r round(naomi_nuts$rhat_max, 3)` (2.5% quantile `r round(naomi_nuts$rhat_lower, 3)`, 50% quantile `r round(naomi_nuts$rhat_median, 3)`, and 97.5% quantile `r round(naomi_nuts$rhat_upper, 3)`; Supplementary Figure 2).
Supplementary Figure 3 shows traceplots for the parameters with the lowest ESS and $\hat R$ respectively.
Based on these diagnostics, we considered the NUTS samples to be of a high quality and suitable for use as a gold-standard.

\begin{reviewer}
\begin{itemize}
\item [R2.2] I really like Figure 5 but just a few more words explaining the green lines would be useful. My first thought was that there was 1 or 3 green lines and that the width meant something. I realise now it's 6561 green lines that either overlap or join together to make thick lines. Perhaps some alpha would help communicate this.
\end{itemize}
\end{reviewer}

\textbf{Response:} We thank the reviewer for the suggestion.
We have adjusted the `alpha` value from 0.5 to 0.1 and added a caption to the figure explaining the possible overlap to make Figure 5 clearer:

```{r}
knitr::include_graphics("revisions/fig5_edit.png")
```

\begin{reviewer}
\begin{itemize}
\item [R2.3] In Figure 6, is the y axis NUTS estimate - GEB estimate? Or the other way around? For the posterior SD, you can interpret positive and negative errors as being conservative or anti-conservative. The new method moves the 3 little clusters at 1.1, 1.4 and 1.8 from negative to positive which seems it could be important and worth noting.
\end{itemize}
\end{reviewer}

Yes that is correct, it's the NUTS estimate minus the GEB or GPCA-AGHQ estimate.
We have clarified the $y$-axis here (and where relevant throughout the manuscript) to read "Absolute difference (NUTS - method)".

\begin{reviewer}
\begin{itemize}
\item [R2.4] Figure 7 is very hard to read. It's an interesting and useful plot. However, I would much prefer the labels to be proper greek characters. And there is just too many colours, so I think it needs to be split into 4 or so groups on seperate panels.
\end{itemize}
\end{reviewer}

\textbf{Response:} We agree that the number of colours in the figure impeded clarity, and have adopted the suggestion to split the figure across panels.
To do this, we have grouped the parameters by the model component (e.g. HIV prevalence), and used colour to represent the type of parameter (e.g. intercept).
The updated version of Figure 7 is shown below:

```{r}
knitr::include_graphics("depends/ks-summary.png")
```

\begin{reviewer}
\begin{itemize}
\item [R2.5] The authors mention that the method is tunable with how many PCA components are selected and what value of k to use. It is a bit of a shame that one or two other options are not examined. But I understand it is already a lot of work to describe all the results. Given that this model is now on the slow side, I would prioritise examining a method that accounts for 80% of the variance, i.e. a faster model rather than a more accurate model. So I will leave it to the authors whether or not to do more work here.
\end{itemize}
\end{reviewer}

\textbf{Response:} In the current work we focused on results which were as accurate as possible while still running in a practical amount of time.
We agree that it is possible that faster results that are nearly as accurate could be obtained using or procedure.

\begin{reviewer}
\begin{itemize}
\item [R2.6] Just looking at Figure 2, a useful intermediate scheme to me seems to be to drop the corners of the grid. So you would be left with 5 points. That said, I think the benefit of this reduces as you get to higher dimensions. So please don't take this comments as work to be done, but rather as an interesting idea for future work.
\end{itemize}
\end{reviewer}

\textbf{Response:} We agree that this could be a useful approach.
This is highlighted in Section 6.1.2., which reads "Finally, it is likely possible to achieve better performance by pruning and prerotation of the quadrature grid, as discussed by @jackel2005note.".
We believe the suggested approach is what we refer to here (very briefly) as "pruning".
To clarify this point, we have rewritten this sentence as:

> Finally, it is likely possible to achieve better performance by pruning (dropping points where the weight is below some threshold) or prerotation (before adaption, in order to e.g. achieve better coverage along the principal axis) of the quadrature grid [@jackel2005note].

# References {#references .unnumbered}